---
title: "Contagion Analysis: Regression Models"
format: html
---

Estimates the baseline LPM and institutional moderation models from the
preregistered analysis plan.

**Specification**:
$$Y_{it} = \beta \cdot \text{NeighborStrike}_{i,t-1:t-2} + \alpha_i + \gamma_{c \times t} + \varepsilon_{it}$$

where $\alpha_i$ are ADM1 fixed effects and $\gamma_{c \times t}$ are
country×week fixed effects. Standard errors clustered at the ADM1 level.

Requires: `data/analysis/contagion_panel.parquet` (from `contagion-data-build.qmd`)

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(lubridate)
library(arrow)
library(fixest)
library(modelsummary)
library(marginaleffects)
library(patchwork)
library(here)

conflicted::conflicts_prefer(dplyr::filter)
conflicted::conflicts_prefer(dplyr::lag)
```

---

## Load data

```{r}
#| label: load-data

panel <- read_parquet(here("data/analysis/contagion_panel.parquet"))

cat("Rows:", nrow(panel), "\n")
cat("ADM1 units:", n_distinct(panel$gid_1), "\n")
cat("Countries:", n_distinct(panel$gid_0), "\n")
cat("Weeks:", n_distinct(panel$week_start), "\n")
```

```{r}
#| label: analysis-sample

# Restrict to observations with non-missing treatment and outcome
# (i.e., regions with at least one within-country neighbor)
analysis <- panel |>
  filter(!is.na(neighbor_strike_within_t1_t2), !is.na(strike_onset))

cat("Analysis sample:", nrow(analysis), "ADM1-weeks\n")
cat("Units dropped (no within-country neighbor or missing outcome):",
    nrow(panel) - nrow(analysis), "\n")
```

---

## Baseline model (H1: demonstration effect)

```{r}
#| label: m-baseline

m_baseline <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1
)

summary(m_baseline)
```

---

## Moderation models (H2–H4)

Each model adds one V-Dem institutional moderator and its interaction with the
treatment. The moderators are standardized (z-scored) so the main effect of
`neighbor_strike_within_t1_t2` can be interpreted at the mean of the moderator.

```{r}
#| label: m-freexp

# H2: Freedom of expression amplifies diffusion
m_freexp <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 * freexp_z | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1
)

summary(m_freexp)
```

```{r}
#| label: m-frassoc

# H3: Freedom of association amplifies diffusion
m_frassoc <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 * frassoc_z | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1
)

summary(m_frassoc)
```

```{r}
#| label: m-repression

# H4: State repression dampens diffusion
m_repression <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 * repression_z | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1
)

summary(m_repression)
```

---

## Regression tables

```{r}
#| label: table-main
#| tbl-cap: "Linear probability models: neighbor strike exposure and institutional moderation"

models <- list(
  "(1) Baseline"   = m_baseline,
  "(2) Expression" = m_freexp,
  "(3) Association" = m_frassoc,
  "(4) Repression"  = m_repression
)

coef_labels <- c(
  "neighbor_strike_within_t1_t2"                  = "Neighbor strike (t-1, t-2)",
  "freexp_z"                                        = "Freedom of expression (z)",
  "frassoc_z"                                       = "Freedom of assoc. (z)",
  "repression_z"                                    = "State repression (z)",
  "neighbor_strike_within_t1_t2:freexp_z"          = "Neighbor strike × Expression",
  "neighbor_strike_within_t1_t2:frassoc_z"         = "Neighbor strike × Association",
  "neighbor_strike_within_t1_t2:repression_z"      = "Neighbor strike × Repression"
)

rows <- tribble(
  ~term,               ~"(1) Baseline", ~"(2) Expression", ~"(3) Association", ~"(4) Repression",
  "ADM1 FE",           "Yes",           "Yes",              "Yes",              "Yes",
  "Country × week FE", "Yes",           "Yes",              "Yes",              "Yes"
)
attr(rows, "position") <- c(10, 11)

modelsummary(
  models,
  coef_map   = coef_labels,
  add_rows   = rows,
  stars      = c("*" = .1, "**" = .05, "***" = .01),
  gof_map    = c("nobs", "r.squared"),
  output     = "gt"
)
```

---

## Marginal effects plots

### H1: Average marginal effect of neighbor strike

```{r}
#| label: plot-ame-baseline
#| fig-width: 5
#| fig-height: 3

avg_slopes(m_baseline, variables = "neighbor_strike_within_t1_t2") |>
  tidy() |>
  ggplot(aes(x = estimate, xmin = conf.low, xmax = conf.high, y = 1)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    title   = "Average marginal effect of neighbor strike",
    x       = "Change in P(strike onset)",
    y       = NULL,
    caption = "LPM with ADM1 and country×week FEs. SEs clustered at ADM1."
  ) +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

### H2–H4: Conditional effects across the distribution of each moderator

```{r}
#| label: plot-conditional-effects
#| fig-width: 10
#| fig-height: 4

plot_conditional <- function(model, moderator, label) {
  plot_slopes(
    model,
    variables  = "neighbor_strike_within_t1_t2",
    condition  = moderator,
    draw       = FALSE
  ) |>
    ggplot(aes(x = .data[[moderator]], y = estimate,
               ymin = conf.low, ymax = conf.high)) +
    geom_ribbon(alpha = 0.15) +
    geom_line() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
    labs(
      title = glue::glue("Conditional effect: {label}"),
      x     = label,
      y     = "Effect of neighbor strike on P(onset)"
    ) +
    theme_minimal()
}

p_freexp    <- plot_conditional(m_freexp,    "freexp_z",    "Freedom of expression (z)")
p_frassoc   <- plot_conditional(m_frassoc,   "frassoc_z",   "Freedom of association (z)")
p_repression <- plot_conditional(m_repression, "repression_z", "State repression (z)")

library(patchwork)
p_freexp | p_frassoc | p_repression
```

---

## Joint moderation model (H2–H4 together, ADM1 + week FEs)

The separate moderation models above use `country×week` FEs, which absorb the
V-Dem moderators (annual, country-level) via collinearity. To identify the
moderator main effects and interactions, we switch to `ADM1 + week` FEs and
include all three moderators simultaneously. This trades some causal
identification of the treatment for the ability to estimate the institutional
interactions.

```{r}
#| label: m-joint-moderation

m_joint_mod <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 * freexp_z +
                 neighbor_strike_within_t1_t2 * frassoc_z +
                 neighbor_strike_within_t1_t2 * repression_z | gid_1 + week_start,
  data    = analysis,
  cluster = ~gid_1
)

summary(m_joint_mod)
```

```{r}
#| label: table-joint-moderation
#| tbl-cap: "Joint moderation model: all three institutional moderators (ADM1 + week FEs)"

joint_coef <- c(
  "neighbor_strike_within_t1_t2"              = "Neighbor strike (t-1, t-2)",
  "freexp_z"                                   = "Freedom of expression (z)",
  "frassoc_z"                                  = "Freedom of assoc. (z)",
  "repression_z"                               = "State repression (z)",
  "neighbor_strike_within_t1_t2:freexp_z"     = "Neighbor strike × Expression",
  "neighbor_strike_within_t1_t2:frassoc_z"    = "Neighbor strike × Association",
  "neighbor_strike_within_t1_t2:repression_z" = "Neighbor strike × Repression"
)

joint_rows <- tribble(
  ~term,               ~"Joint model",
  "ADM1 FE",           "Yes",
  "Week FE",           "Yes"
)
attr(joint_rows, "position") <- c(10, 11)

modelsummary(
  list("Joint model" = m_joint_mod),
  coef_map   = joint_coef,
  add_rows   = joint_rows,
  stars      = c("*" = .1, "**" = .05, "***" = .01),
  gof_map    = c("nobs", "r.squared"),
  output     = "gt"
)
```

---

## Robustness checks

### Separate t-1 and t-2 lags

```{r}
#| label: robustness-separate-lags

m_lag1 <- feols(
  strike_onset ~ neighbor_strike_within_t1_t1 | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1
)

m_lag2 <- feols(
  strike_onset ~ neighbor_strike_within_t2_t2 | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1
)

modelsummary(
  list("t-1 only" = m_lag1, "t-2 only" = m_lag2, "Pooled t-1:t-2" = m_baseline),
  coef_map = c(
    "neighbor_strike_within_t1_t1" = "Neighbor strike (t-1)",
    "neighbor_strike_within_t2_t2" = "Neighbor strike (t-2)",
    "neighbor_strike_within_t1_t2" = "Neighbor strike (pooled)"
  ),
  stars  = c("*" = .1, "**" = .05, "***" = .01),
  gof_map = c("nobs", "r.squared"),
  output = "gt"
)
```

### Logistic regression

```{r}
#| label: robustness-logit

m_logit <- feglm(
  strike_onset ~ neighbor_strike_within_t1_t2 | gid_1 + country_week,
  family  = binomial,
  data    = analysis,
  cluster = ~gid_1
)

summary(m_logit)
```

### Two-way clustering (ADM1 + country-week)

```{r}
#| label: robustness-twoway-cluster

m_twoway <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 | gid_1 + country_week,
  data    = analysis,
  cluster = ~gid_1 + country_week
)

modelsummary(
  list("ADM1 cluster" = m_baseline, "Two-way cluster" = m_twoway),
  coef_map = c("neighbor_strike_within_t1_t2" = "Neighbor strike (t-1, t-2)"),
  stars    = c("*" = .1, "**" = .05, "***" = .01),
  gof_map  = c("nobs", "r.squared"),
  output   = "gt"
)
```

---

## National vs. neighborhood contagion decomposition

The baseline LPM with `country×week` FEs yields a null treatment effect, while
`ADM1 + week` FEs yield a positive effect (~1.5 pp). To understand why, we
decompose the apparent neighborhood effect into:

- **Country-level contagion**: did any other region in the same country strike
  in $t-1$ or $t-2$? (`country_strike_t1_t2`)
- **Neighborhood-specific diffusion**: does a geographic neighbor's strike
  predict onset *above and beyond* the country-level signal?

```{r}
#| label: build-country-strike

# Country-week total strike count, then lagged
country_weekly <- panel |>
  group_by(gid_0, week_start) |>
  summarize(country_strike_count = sum(strike_onset), .groups = "drop")

panel_decomp <- panel |>
  left_join(country_weekly, by = c("gid_0", "week_start")) |>
  arrange(gid_0, week_start) |>
  group_by(gid_0) |>
  mutate(
    country_strike_t1    = dplyr::lag(country_strike_count, 1),
    country_strike_t2    = dplyr::lag(country_strike_count, 2),
    country_strike_t1_t2 = as.integer((country_strike_t1 + country_strike_t2) > 0)
  ) |>
  ungroup()

analysis_decomp <- panel_decomp |>
  filter(
    !is.na(neighbor_strike_within_t1_t2),
    !is.na(strike_onset),
    !is.na(country_strike_t1_t2)
  )

cat("Decomposition sample:", nrow(analysis_decomp), "ADM1-weeks\n")
cat("Country strike prevalence:",
    round(mean(analysis_decomp$country_strike_t1_t2) * 100, 1), "%\n")
```

```{r}
#| label: m-country-only

# M1: country-level contagion only
m_country <- feols(
  strike_onset ~ country_strike_t1_t2 | gid_1 + week_start,
  data    = analysis_decomp,
  cluster = ~gid_1
)

summary(m_country)
```

```{r}
#| label: m-neighbor-only-adm1week

# M2: neighborhood contagion only (ADM1 + week FEs for comparability)
m_neighbor_aw <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 | gid_1 + week_start,
  data    = analysis_decomp,
  cluster = ~gid_1
)

summary(m_neighbor_aw)
```

```{r}
#| label: m-decomp

# M3: both together — does geographic neighbor add anything beyond country wave?
m_decomp <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 + country_strike_t1_t2 |
    gid_1 + week_start,
  data    = analysis_decomp,
  cluster = ~gid_1
)

summary(m_decomp)
```

```{r}
#| label: m-decomp-frassoc

# M4: add frassoc moderation on the neighborhood term
m_decomp_frassoc <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 * frassoc_z +
                 country_strike_t1_t2 | gid_1 + week_start,
  data    = analysis_decomp,
  cluster = ~gid_1
)

summary(m_decomp_frassoc)
```

```{r}
#| label: table-decomp
#| tbl-cap: "National vs. neighborhood contagion decomposition"

decomp_models <- list(
  "(1) Country"        = m_country,
  "(2) Neighbor"       = m_neighbor_aw,
  "(3) Decomposition"  = m_decomp,
  "(4) + frassoc"      = m_decomp_frassoc
)

decomp_coef <- c(
  "neighbor_strike_within_t1_t2"         = "Neighbor strike (t-1, t-2)",
  "country_strike_t1_t2"                 = "Country strike (t-1, t-2)",
  "frassoc_z"                            = "Freedom of assoc. (z)",
  "neighbor_strike_within_t1_t2:frassoc_z" = "Neighbor strike × Assoc."
)

decomp_rows <- tribble(
  ~term,               ~"(1) Country", ~"(2) Neighbor", ~"(3) Decomposition", ~"(4) + frassoc",
  "ADM1 FE",           "Yes",          "Yes",            "Yes",                "Yes",
  "Week FE",           "Yes",          "Yes",            "Yes",                "Yes"
)
attr(decomp_rows, "position") <- c(8, 9)

modelsummary(
  decomp_models,
  coef_map   = decomp_coef,
  add_rows   = decomp_rows,
  stars      = c("*" = .1, "**" = .05, "***" = .01),
  gof_map    = c("nobs", "r.squared"),
  output     = "gt"
)
```

---

## Exploratory: cross-border diffusion

Sample restricted to ADM1 units with at least one cross-border neighbor (~95
units). `country×week` FEs are appropriate here: the treatment is a *foreign*
neighbor's behavior, so it is not absorbed by the focal region's country-week
FE. Both `country×week` and `ADM1 + week` specs are shown for comparability.

```{r}
#| label: exploratory-cross-border

cross_data <- analysis |> filter(!is.na(neighbor_strike_cross_t1_t2))
cat("Cross-border sample: ADM1 units =", n_distinct(cross_data$gid_1),
    "| rows =", nrow(cross_data), "\n")

# Country×week FE (treatment is foreign, FE controls own-country shocks)
m_cross_cw <- feols(
  strike_onset ~ neighbor_strike_cross_t1_t2 | gid_1 + country_week,
  data    = cross_data,
  cluster = ~gid_1
)

# ADM1 + week FE (for comparability with main specs)
m_cross_aw <- feols(
  strike_onset ~ neighbor_strike_cross_t1_t2 | gid_1 + week_start,
  data    = cross_data,
  cluster = ~gid_1
)

summary(m_cross_cw)
summary(m_cross_aw)
```

```{r}
#| label: exploratory-joint

joint_data <- analysis |>
  filter(!is.na(neighbor_strike_within_t1_t2),
         !is.na(neighbor_strike_cross_t1_t2))

# Within + cross-border, country×week FE
m_joint_cw <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 + neighbor_strike_cross_t1_t2 |
    gid_1 + country_week,
  data    = joint_data,
  cluster = ~gid_1
)

# Within + cross-border, ADM1 + week FE
m_joint_aw <- feols(
  strike_onset ~ neighbor_strike_within_t1_t2 + neighbor_strike_cross_t1_t2 |
    gid_1 + week_start,
  data    = joint_data,
  cluster = ~gid_1
)

modelsummary(
  list(
    "Cross only (cw FE)"    = m_cross_cw,
    "Cross only (week FE)"  = m_cross_aw,
    "Joint (cw FE)"         = m_joint_cw,
    "Joint (week FE)"       = m_joint_aw
  ),
  coef_map = c(
    "neighbor_strike_within_t1_t2" = "Neighbor strike within (t-1, t-2)",
    "neighbor_strike_cross_t1_t2"  = "Neighbor strike cross-border (t-1, t-2)"
  ),
  stars   = c("*" = .1, "**" = .05, "***" = .01),
  gof_map = c("nobs", "r.squared"),
  output  = "gt"
)
```

---

## Validated locations only

Robustness check restricting to events where the ADM1 geocode was confirmed
correct by the LLM validation step (`geo_quality == "validated"`). Unvalidated
strike weeks are recoded as zero-strike weeks, keeping the balanced panel intact.

```{r}
#| label: validated-sample

panel_val <- panel |>
  mutate(
    strike_onset_val = as.integer(geo_quality == "validated")
  )

panel_val_decomp <- panel_decomp |>
  mutate(
    strike_onset_val = as.integer(geo_quality == "validated")
  )

analysis_val <- panel_val |>
  filter(!is.na(neighbor_strike_within_t1_t2))

analysis_val_decomp <- panel_val_decomp |>
  filter(!is.na(neighbor_strike_within_t1_t2),
         !is.na(country_strike_t1_t2))

cat("Validated outcome prevalence:",
    round(mean(analysis_val$strike_onset_val) * 100, 2), "%\n")
cat("(vs.", round(mean(analysis$strike_onset) * 100, 2),
    "% in full sample)\n")
```

```{r}
#| label: validated-baseline

m_val_baseline <- feols(
  strike_onset_val ~ neighbor_strike_within_t1_t2 | gid_1 + country_week,
  data    = analysis_val,
  cluster = ~gid_1
)

m_val_aw <- feols(
  strike_onset_val ~ neighbor_strike_within_t1_t2 | gid_1 + week_start,
  data    = analysis_val,
  cluster = ~gid_1
)

summary(m_val_baseline)
summary(m_val_aw)
```

```{r}
#| label: validated-decomp

m_val_decomp <- feols(
  strike_onset_val ~ neighbor_strike_within_t1_t2 + country_strike_t1_t2 |
    gid_1 + week_start,
  data    = analysis_val_decomp,
  cluster = ~gid_1
)

modelsummary(
  list(
    "Full sample\n(country×week)" = m_baseline,
    "Full sample\n(week FE)"      = m_val_aw,
    "Validated only\n(country×week)" = m_val_baseline,
    "Validated only\n(week FE)"   = m_val_aw,
    "Validated: decomp"           = m_val_decomp
  ),
  coef_map = c(
    "neighbor_strike_within_t1_t2" = "Neighbor strike (t-1, t-2)",
    "country_strike_t1_t2"         = "Country strike (t-1, t-2)"
  ),
  stars   = c("*" = .1, "**" = .05, "***" = .01),
  gof_map = c("nobs", "r.squared"),
  output  = "gt"
)
```
