---
title: "Filter GDELT Events by Themes and URLS"
format: html
---

# Filter by Themes

## Setup

```{r}
#| label: setup

library(tidyverse)
library(arrow)
```

## Create a List of Themes

```{r}
# Load the enriched GDELT dataset with themes
gdelt_strikes_enriched <- read_parquet("data/enhanced/gdelt_strikes_enriched.parquet")

# Explode themes into long format, drop offsets
themes_long <- gdelt_strikes_enriched |>
  select(GLOBALEVENTID, all_themes) |>
  filter(!is.na(all_themes)) |>
  separate_rows(all_themes, sep = ";") |>
  mutate(all_themes = str_trim(all_themes)) |>
  filter(all_themes != "") |>
  mutate(theme = sub(",.*$", "", all_themes)) |>
  distinct(GLOBALEVENTID, theme) |>
  select(GLOBALEVENTID, theme)

# Frequency table of unique themes
theme_counts <- themes_long |>
  count(theme, sort = TRUE)

print(theme_counts, n = 100)

# Export to CSV for review
write_csv(theme_counts, "data/gdelt-themes/theme_counts.csv")
```


There are 12,113 themes in total. Let's try filtering first with some obvious keywords.

```{r}
# Candidate strike-related themes (first-pass via keywords)
candidate_themes <- theme_counts %>%
  filter(
    str_detect(theme,
      regex("STRIKE|WALKOUT|LOCKOUT|PICKET|PROTEST|DEMONSTRAT|UNION|BOYCOTT|UNREST|CONTRACT_DISPUTE",
            ignore_case = TRUE)
    )
  ) %>%
  filter(
    !str_detect(theme, regex("DRONE_STRIKE", ignore_case = TRUE)),
    !str_detect(theme, regex("^TAX_POLITICAL_PARTY_", ignore_case = TRUE)),
    !str_detect(theme, regex("^TAX_TERROR_GROUP_", ignore_case = TRUE))
  )
```

Step 1: Narrow down to strikes only by simply filtering for .

```{r}
# --- Define lists of inclusions / exclusions ---
entry_terms <- c("STRIKE", "WALKOUT", "LOCKOUT", "PICKET", "UNION", "CONTRACT_DISPUTE")

exclude_terms <- c("DRONE_STRIKE", "RELIGION", "PROTESTANT", "FREETRADE", 
                   "CUSTOMS", "CREDIT", "HUNGERSTRIKE", "CONTRACT_DISPUTE", 
                   "TEACHER")

# --- Apply filtering ---
strike_themes <- theme_counts |>
  filter(str_detect(theme, regex(paste(entry_terms, collapse = "|"), ignore_case = TRUE))) |>
  filter(!str_detect(theme, regex(paste(exclude_terms, collapse = "|"), ignore_case = TRUE))) |>
  filter(!str_detect(theme, regex("^TAX_POLITICAL_PARTY_", ignore_case = TRUE))) |>
  filter(!str_detect(theme, regex("^TAX_TERROR_GROUP_", ignore_case = TRUE))) |>
  filter(!str_detect(theme, regex("^TAX_FNCACT_.*WORKER", ignore_case = TRUE)))   # keep workers for secondary use
```

```{r}
# econdary analysis set — extract the worker categories separately
worker_categories <- theme_counts %>%
  filter(str_detect(theme, regex("^TAX_FNCACT_.*WORKER", ignore_case = TRUE)))
```

Step 2: Code for who was striking



Step 3: Code for why the strike took place or issues that were at stake



Step 4: Code for economic sectors affected



Setp 5: Code for the nature of protest, e.g. violence, peaceful, etc. 



## Create Strikes Dataset

```{r}
gdelt_strikes <- gdelt_strikes_enriched |>
    filter(str_detect(all_themes, "STRIKE"))

# Save to Parquet
write_parquet(gdelt_strikes, "data/enhanced/gdelt_strikes.parquet")
```


## Create Labor-Related Protest Dataset