---
title: "Forecasting Strikes with GDELT Event Data"
subtitle: "A Preliminary Analysis"
author: "Emmanuel Teitelbaum"
date: today
date-format: long
footer: "[View Our Code and Data](https://github.com/eteitelbaum/code-satp)"
logo: images/esia-logo.png
format:
  revealjs:
    theme: [simple, custom.scss]
    transition: fade
    slide-number: true
    chalkboard: true
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

# Packages
library(arrow)
library(dplyr)
library(stringr)
library(janitor)
library(lubridate)
library(readr)
library(ggplot2)
library(kableExtra)
library(broom)

# Set seed for reproducible sampling
set.seed(74)

# Root-aware paths via here(); avoids ../ from subfolders
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
root <- here::here()
paths <- list(
  raw              = file.path(root, "data/raw/gdelt_strikes.parquet"),
  filtered         = file.path(root, "data/raw/gdelt_strikes_filtered.parquet"),
  adm_week_full    = file.path(root, "data/analysis/adm_week_full.parquet"),
  hiconf           = file.path(root, "data/raw/gdelt_strikes_hiconf.parquet"),
  gkg_event_level  = file.path(root, "data/raw/strike_event_gkg.parquet"),
  enriched         = file.path(root, "data/enhanced/gdelt_strikes_enriched.parquet"),
  final_strikes    = file.path(root, "data/enhanced/gdelt_strikes.parquet")
)

# Source helpers kept in repo
helpers_path <- file.path(root, "helpers/helpers.R")
if (file.exists(helpers_path)) source(helpers_path)

# Preload only datasets needed for illustrative URL samples
adm1_only  <- arrow::read_parquet(paths$filtered)
hiconf     <- arrow::read_parquet(paths$hiconf)  
seg        <- arrow::read_parquet(paths$gkg_event_level)
final      <- arrow::read_parquet(paths$final_strikes)
```

# Part 1: Mining GDELT 

## What is GDELT? 

<br>

- Global news event database coded with the CAMEO taxonomy.
- Strikes fall under Protest (EventRootCode 14); EventCode 143 covers Strike/Boycott.
- Subtypes commonly used include: 1431 Work Stoppage, 1432 General Strike, 1433 Boycott.
- In this project we targeted EventCode 143 and its subcategories.

## Google BigQuery (public datasets)

<br>

- Serverless data warehouse; SQL at scale with on-demand pricing.
- GDELT is hosted as public datasets under `gdelt-bq.gdeltv2`.
- We query public tables and materialize intermediate results in our project for reuse.

## GDELT tables used in this study

<br>

- `events`: event-level records (geography, date, actors, intensity, tone, source URL).
- `eventmentions`: article mentions of events (Confidence, source type, tone).
- `gkg`: Global Knowledge Graph (per-document themes, tone, people, source URLs).

## Pipeline overview

<div style="text-align: center;">
```{mermaid}
graph TD
    A[Raw events] --> B[ADM1 filter]
    B --> C[High-confidence filter]
    C --> D[GKG enrichment]
    D --> E[STRIKE theme filter]
    E --> F[Final dataset]
```
</div>

## Step 1: Raw events extraction

<br> 

Downloaded strike events with latitude and longitude info. Yields table with **492,500 rows**. 

```sql
SELECT *
FROM `gdelt-bq.gdeltv2.events`
WHERE EventCode LIKE '143%'
  AND SQLDATE >= 20150101
  AND ActionGeo_Lat IS NOT NULL
  AND ActionGeo_Long IS NOT NULL;
```

## Step 2: ADM1 geographic filter

<br>

Filtered for events with valid ADM1 codes. Yields table with **328,982 rows**.

```{r}
#| label: adm1-filter-save
#| echo: true
#| eval: false

adm1_only <- raw |>
  dplyr::filter(
    !is.na(ActionGeo_ADM1Code),
    !is.na(ActionGeo_CountryCode),
    ActionGeo_ADM1Code != ActionGeo_CountryCode
  )

arrow::write_parquet(
  adm1_only,
  file.path(root, "data/raw/gdelt_strikes_filtered.parquet")
)
```

## Step 3: High-confidence filter

Queried the `eventmentions` table to find events with high-confidence mentions (Confidence ≥80). Leaves us with **134,141 events.**

```sql
WITH high_conf AS (
  SELECT m.GLOBALEVENTID,
         COUNTIF(m.Confidence >= 80) AS high_conf_mentions,
         COUNT(DISTINCT m.MentionSourceName) AS unique_sources
  FROM `gdelt-bq.gdeltv2.eventmentions` m
  JOIN `gdelt-449818.gdelt_analysis.strike_events` s
    ON m.GLOBALEVENTID = s.GLOBALEVENTID
  GROUP BY m.GLOBALEVENTID
)
SELECT s.GLOBALEVENTID, h.high_conf_mentions, h.unique_sources
FROM `gdelt-449818.gdelt_analysis.strike_events` s
JOIN high_conf h ON s.GLOBALEVENTID = h.GLOBALEVENTID
WHERE h.high_conf_mentions > 0;
```

## Step 4: GKG enrichment pipeline

<br>

- This was a multi-step process to enrich strike events with tone metrics and themes from news articles.
- Goal was to link tone, themes and URLs from the GDELT Knowledge Graph (GKG) to the strikes in the events table
- To do that we first had to filter the documents related to each strike from the `eventmentions` table

## Step 4.1: Extract mention identifiers

<br>
Select only eventmentions that refer to high-confidence strike events.

```sql
CREATE OR REPLACE TABLE `gdelt-449818.gdelt_analysis.strike_mentions` AS
SELECT DISTINCT m.MentionIdentifier, m.GLOBALEVENTID
FROM `gdelt-bq.gdeltv2.eventmentions` m
JOIN `gdelt-449818.gdelt_analysis.strike_events_hiconf` s
  ON m.GLOBALEVENTID = s.GLOBALEVENTID;
```

## Step 4.2: Filter GKG documents

<br>
Materialize only GKG documents that mention strike events. **Expensive scan (~20TB, ~$128): Do once.**

```sql
CREATE OR REPLACE TABLE `gdelt-449818.gdelt_analysis.strike_gkg_docs` AS
SELECT g.*
FROM `gdelt-bq.gdeltv2.gkg` g
JOIN `gdelt-449818.gdelt_analysis.strike_mentions` sm
  ON g.DocumentIdentifier = sm.MentionIdentifier
WHERE g.DATE >= 20150101;
```

## Step 4.3: Aggregate tone and URLs

<br>
Aggregate document-level tone to event level and collect article URLs.

```sql
CREATE OR REPLACE TABLE gdelt-449818.gdelt_analysis.strike_event_tone AS
SELECT m.GLOBALEVENTID,
       AVG(CAST(SPLIT(g.V2Tone, ',')[OFFSET(0)] AS FLOAT64)) AS avg_tone,
       ARRAY_AGG(DISTINCT g.DocumentIdentifier IGNORE NULLS) AS article_urls
FROM `gdelt-449818.gdelt_analysis.strike_mentions` m
JOIN `gdelt-449818.gdelt_analysis.strike_gkg_docs` g
  ON m.MentionIdentifier = g.DocumentIdentifier
GROUP BY m.GLOBALEVENTID;
```

## Step 4.4: Aggregate themes

<br>
Tokenize GKG V2Themes and aggregate unique themes per event.

```sql
-- Event-level themes (tokenize + aggregate)
CREATE OR REPLACE TABLE gdelt-449818.gdelt_analysis.strike_event_themes AS
SELECT m.GLOBALEVENTID,
       STRING_AGG(DISTINCT TRIM(theme), ';') AS all_themes
FROM `gdelt-449818.gdelt_analysis.strike_mentions` m
JOIN `gdelt-449818.gdelt_analysis.strike_gkg_docs` g
  ON m.MentionIdentifier = g.DocumentIdentifier
CROSS JOIN UNNEST(SPLIT(g.V2Themes, ';')) AS theme
GROUP BY m.GLOBALEVENTID;
```

## Step 4.5: Combine GKG enrichments

<br>
Combine tone and themes into single reusable event-level table.

```sql
-- Final event-level enrichment table
CREATE OR REPLACE TABLE gdelt-449818.gdelt_analysis.strike_event_gkg AS
SELECT t.*, th.all_themes
FROM gdelt-449818.gdelt_analysis.strike_event_tone t
LEFT JOIN gdelt-449818.gdelt_analysis.strike_event_themes th
  ON t.GLOBALEVENTID = th.GLOBALEVENTID;
```

## What GKG enrichments are used for

**Tone metrics** → Used as predictors in strike forecasting models, measure overall sentiment of news coverage (scale: -100 to +100, but typically ranges -10 to +10)

**Mention/Article counts** → Captures media attention and event significance, e.g. total event mentions, total articles related to event

**Themes** → Used to filter events to strike-relevant content (STRIKE theme filter)

## Theme exploration

| Example themes found in our events (before filtering):
- STRIKE (direct match)
- PROTEST (related activity)
- TAX_FNCACT_WORKERS, TAX_FNCACT_EMPLOYEE (labor-related)
- ECON_UNIONS (union activity)
- WB_2811_COLLECTIVE_BARGAINING (negotiation context)

## Why we chose "STRIKE" theme only

- We filtered for events where themes contain "STRIKE" because:
  - Most specific indicator of strike events
  - Balances precision (true strikes) vs. recall (catching most strikes)
  - Other themes too broad (PROTEST) or too narrow (specific union codes)
- Left us with **30,899 strike events**.

## Data quality validation

```{r}
#| label: validation-plot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

# Load coded results
url_coded <- read_csv("url_validation_coded.csv")

# Calculate precision by stage
validation_summary <- url_coded |>
  group_by(stage) |>
  summarise(
    n_urls = n(),
    n_relevant = sum(is_relevant),
    precision = round(mean(is_relevant) * 100, 1)
  ) |>
  mutate(
    stage_label = case_when(
      stage == "1_ADM1" ~ "ADM1 filter",
      stage == "2_HighConf" ~ "+ High-confidence",
      stage == "3_STRIKE" ~ "+ STRIKE theme"
    ),
    stage_order = case_when(
      stage == "1_ADM1" ~ 1,
      stage == "2_HighConf" ~ 2,
      stage == "3_STRIKE" ~ 3
    )
  ) |>
  arrange(stage_order)

# Create bar chart
ggplot(validation_summary, aes(x = reorder(stage_label, stage_order), y = precision)) +
  geom_col(fill = "steelblue", alpha = 0.8) +
  geom_text(aes(label = paste0(precision, "%")), vjust = -0.5, size = 4) +
  labs(
    title = "Filtering Effectiveness",
    subtitle = "Percentage of URLs actually about strikes at each filtering stage",
    x = "Filter Stage",
    y = "Precision (% Strike-Relevant URLs)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  ) +
  ylim(0, max(validation_summary$precision) * 1.1)
```

## Examples: Definite strikes {.smaller}

```{r}
#| label: definite-strikes
#| echo: false
#| message: false
#| warning: false

# Load coded results and get definite strikes
url_coded <- read_csv("url_validation_coded.csv")

strikes <- url_coded |> 
  filter(is_relevant == 1) |> 
  slice(2:5) |>  # Skip Portugal trains, get more obvious strikes
  select(url, notes)

# Display as simple table
knitr::kable(strikes, 
             col.names = c("URL", "Description"),
             format = "html") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"))
```

## Examples: Definite non-strikes {.smaller}

```{r}
#| label: definite-non-strikes
#| echo: false
#| message: false
#| warning: false

# Get definite non-strikes
non_strikes <- url_coded |> 
  filter(is_relevant == 0) |> 
  head(4) |>
  select(url, notes)

# Display as simple table
knitr::kable(non_strikes, 
             col.names = c("URL", "Description"),
             format = "html") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"))
```

## Examples: Borderline cases {.smaller}

```{r}
#| label: borderline-cases
#| echo: false
#| message: false
#| warning: false

# Get some borderline cases (mix of relevant/not relevant)
# Include Portugal train example (less obvious strike) + two from STRIKE stage
portugal_train <- url_coded |> 
  filter(is_relevant == 1) |> 
  slice(1) |>
  select(url, notes)

strike_stage_examples <- url_coded |> 
  filter(stage == "3_STRIKE") |>  # After STRIKE filter
  arrange(desc(is_relevant)) |>   # Put relevant first
  slice(1:2) |>  # Get first two examples
  select(url, notes)

borderline <- bind_rows(strike_stage_examples, portugal_train) |>
  mutate(
    # Break up the long Mexican URL for better readability
    url = case_when(
      str_detect(url, "noticias.terra.es") ~ 
        "https://noticias.terra.es/mundo/latinoamerica/\nmaestros-marchan-en-la-ciudad-de-mexico\n-en-contra-de-la-reforma-educativa",
      TRUE ~ url
    )
  )

# Display as simple table
knitr::kable(borderline, 
             col.names = c("URL", "Description"),
             format = "html") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"))
```

# Part 2: Data Exploration

```{r}
#| label: part2-setup

# Load the panel dataset for analysis
library(tidyverse)
library(ggplot2)
library(scales)
library(viridis)

# Load adm_week_full data
adm_week_full <- arrow::read_parquet(file.path(root, "data/analysis/adm_week_full.parquet"))

# Basic dataset info
# cat("Panel dataset dimensions:", nrow(adm_week_full), "region-weeks\n")
# cat("Unique ADM1 regions:", n_distinct(adm_week_full$action_geo_adm1_code), "\n")
# cat("Date range:", min(adm_week_full$date), "to", max(adm_week_full$date), "\n")
# cat("Total weeks:", n_distinct(adm_week_full$date), "\n")
```

## From events to panel data

- **30,899 strike events** → **806,000 region-weeks** 
- **1,550 ADM1 regions** × **560 weeks** = full panel
- **Target**: Weekly strike counts per region
- **Challenge**: Most region-weeks have zero strikes

```{r}
#| label: panel-structure

# Calculate panel structure summary
panel_summary <- adm_week_full |>
  summarise(
    total_observations = n(),
    unique_regions = n_distinct(action_geo_adm1_code),
    unique_countries = n_distinct(action_geo_country_code),
    total_weeks = n_distinct(date),
    date_range = paste(min(date), "to", max(date))
  )

# Display as formatted text
# cat("Panel Structure:\n")
# cat("- Total observations:", format(panel_summary$total_observations, big.mark = ","), "region-weeks\n")
# cat("- Unique ADM1 regions:", format(panel_summary$unique_regions, big.mark = ","), "\n")
# cat("- Unique countries:", format(panel_summary$unique_countries, big.mark = ","), "\n")
# cat("- Time span:", panel_summary$total_weeks, "weeks (", round(panel_summary$total_weeks/52, 1), "years)\n")
# cat("- Date range:", panel_summary$date_range, "\n")
```

## The Zero-inflation problem

```{r}
#| label: zero-inflation-analysis
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

# Calculate zero-inflation and overdispersion statistics
zero_stats <- adm_week_full |>
  summarise(
    # Zero inflation
    zero_count = sum(strike_count == 0),
    positive_count = sum(strike_count > 0),
    zero_rate = mean(strike_count == 0),
    positive_rate = mean(strike_count > 0),
    
    # Distribution moments
    mean_count = mean(strike_count),
    var_count = var(strike_count),
    overdispersion_ratio = var_count / mean_count,
    
    # Quantiles
    p50 = quantile(strike_count, 0.5),
    p90 = quantile(strike_count, 0.9),
    p95 = quantile(strike_count, 0.95),
    p99 = quantile(strike_count, 0.99),
    max_count = max(strike_count)
  )

# Create zero vs positive weeks bar chart
zero_positive_data <- data.frame(
  category = c("Zero Strikes", "Positive Strikes"),
  count = c(zero_stats$zero_count, zero_stats$positive_count),
  percentage = c(zero_stats$zero_rate * 100, zero_stats$positive_rate * 100)
)

# Create the visualization with key statistics in caption
ggplot(zero_positive_data, aes(x = category, y = percentage, fill = category)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            vjust = -0.5, size = 5, fontface = "bold") +
  scale_fill_manual(values = c("Zero Strikes" = viridis(10)[2], 
                               "Positive Strikes" = viridis(10)[8])) +
  labs(
    title = "Distribution of Strike Activity",
    subtitle = paste0("Mean: ", round(zero_stats$mean_count, 3), " strikes/week | ",
                      "Overdispersion ratio: ", round(zero_stats$overdispersion_ratio, 1), 
                      " (variance >> mean)"),
    x = "Strike Count",
    y = "Percentage of Region-Weeks"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11)
  ) +
  ylim(0, max(zero_positive_data$percentage) * 1.1)
```

## Strike counts

```{r}
#| label: positive-distribution-plot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

# Create histogram of positive counts (capped at reasonable range for visualization)
positive_data <- adm_week_full |>
  filter(strike_count > 0) |>
  mutate(
    strike_count_capped = pmin(strike_count, 10)  # Cap at 10 for better visualization
  )

# Calculate summary statistics for the note
positive_mean <- mean(adm_week_full$strike_count[adm_week_full$strike_count > 0])
positive_median <- median(adm_week_full$strike_count[adm_week_full$strike_count > 0])
positive_total <- sum(adm_week_full$strike_count > 0)

ggplot(positive_data, aes(x = strike_count_capped)) +
  geom_histogram(bins = 10, alpha = 0.8, fill = viridis(10)[8]) +
  scale_x_continuous(breaks = 1:10, labels = c(1:9, "10+")) +
  labs(
    title = "Distribution of Positive Strike Counts",
    subtitle = paste0(format(positive_total, big.mark = ","), " region-weeks with strikes > 0"),
    x = "Strike Count",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11)
  )
```

## Geographic concentration

```{r}
#| label: geographic-concentration-plot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

# Top regions by total strike activity
top_regions <- adm_week_full |>
  group_by(action_geo_adm1_code, action_geo_country_code) |>
  summarise(
    total_strikes = sum(strike_count),
    positive_weeks = sum(strike_count > 0),
    .groups = "drop"
  ) |>
  arrange(desc(total_strikes)) |>
  head(15)

# Calculate concentration metrics
total_strikes_all <- sum(adm_week_full$strike_count)
top_15_strikes <- sum(top_regions$total_strikes)
concentration_pct <- round(top_15_strikes / total_strikes_all * 100, 1)

# Plot top regions by strike activity
top_regions_plot <- top_regions |>
  head(10) |>
  mutate(
    region_label = paste0(action_geo_country_code, "-", str_sub(action_geo_adm1_code, 1, 8)),
    region_label = fct_reorder(region_label, total_strikes)
  )

ggplot(top_regions_plot, aes(x = total_strikes, y = region_label)) +
  geom_col(alpha = 0.8, fill = viridis(10)[3]) +
  labs(
    title = "Top 10 Regions by Total Strike Activity",
    subtitle = paste0("Top 15 regions account for ", concentration_pct, "% of all strikes"),
    x = "Total Strikes",
    y = "Region (Country-ADM1)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

## Time trends

```{r}
#| label: temporal-concentration-plot
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

# Aggregate strikes by week for time series
weekly_totals <- adm_week_full |>
  group_by(date) |>
  summarise(
    total_strikes = sum(strike_count),
    regions_with_strikes = sum(strike_count > 0),
    .groups = "drop"
  )

# Calculate temporal statistics
mean_weekly_strikes <- mean(weekly_totals$total_strikes)
cv_weekly_strikes <- sd(weekly_totals$total_strikes) / mean_weekly_strikes
peak_week_date <- weekly_totals$date[which.max(weekly_totals$total_strikes)]
peak_week_strikes <- max(weekly_totals$total_strikes)

# Time series plot of weekly strike totals
ggplot(weekly_totals, aes(x = date, y = total_strikes)) +
  geom_line(color = viridis(10)[3], alpha = 0.8) +
  geom_smooth(method = "loess", se = FALSE, color = viridis(10)[8], linewidth = 1.2) +
  labs(
    title = "Global Strike Activity Over Time",
    subtitle = paste0("Mean: ", round(mean_weekly_strikes, 1), " strikes/week | ",
                      "CV: ", round(cv_weekly_strikes, 2), " (high variability)"),
    x = "Date",
    y = "Total Weekly Strikes"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11)
  )
```

## Temporal stats

```{r}
# Create temporal statistics table
temporal_stats <- tibble(
  Metric = c("Mean strikes per week", "Coefficient of variation", "Peak week", "Peak week strikes"),
  Value = c(
    paste0(round(mean_weekly_strikes, 1), " strikes"),
    paste0(round(cv_weekly_strikes, 2), " (high variability)"),
    as.character(peak_week_date),
    paste0(peak_week_strikes, " strikes")
  )
)

kable(temporal_stats)
```

## Why this matters for modeling {.smaller}

- **Zero-inflation (98% zeros)** → Need distributions that handle excess zeros (Tweedie, hurdle models)
- **Overdispersion (variance >> mean)** → Can't use plain Poisson regression  
- **Geographic concentration** → Need spatial features and regional fixed effects
- **Temporal clustering** → Need lagged features and time trends
- **Panel structure** → Need models that account for repeated observations per region

# Part 3: Modeling Strategy

## Modeling rare events

**Data characteristics:**

- Zero-inflated count data 
- Overdispersion (variance >> mean)
- Panel structure: 1,550 ADM1 regions 

**Task:** Predict weekly strike counts at the ADM1 geographic level

## Modeling approaches: Single-stage

**Single-stage models** predict counts directly in one step:

- **Native count distributions**: Poisson, Tweedie GLMs
- **Log-transformed targets**: log1p(count) + backtransform
  - Linear model (LM)
  - Elastic net (GLMNET)
  - Gradient boosting (LightGBM, XGBoost)

## Modeling approaches: Two-stage (hurdle) {.smaller}

**Two-stage (hurdle) models** decompose prediction into:

1. **Stage 1 (Occurrence)**: Classify P(strike > 0)
   - Logistic regression (GLM, GLMNET)
   - Tree-based classifiers (LightGBM, XGBoost)

2. **Stage 2 (Severity)**: Predict count given occurrence
   - Gamma regression (LightGBM, XGBoost)
   - Log-transformed models (LM, GLMNET, LightGBM, XGBoost)

3. **Combine**: E[count] = P(occurrence) × E[severity | occurrence]

## Features: Temporal {.smaller}

**Seasonality and trends:**

- Calendar effects: Month, quarter, and year-end indicators
- Long-term patterns: 10-year cyclical trends encoded via sine/cosine transformations
- Linear time trend to capture overall trajectory

**Lagged strike activity:**

- Recent history: Strike counts from 1, 2, and 4 weeks prior
- Smoothed trends: Rolling averages over 4 and 8 weeks
- Capture strong temporal autocorrelation in strike occurrence

## Features: Media and actors {.smaller}

**Media coverage (lagged):**

- Volume: Total news articles and event mentions
- Sentiment: Average tone of news coverage
- Capture media attention and public sentiment around labor activity

**Actor characteristics:**

- Diversity: Number of unique actor types involved
- Composition: Proportions of government, labor, and civil society actors
- Measure coordination and coalition formation

## Features: Spatial and geographic {.smaller}

**Spatial spillovers:**

- Country-level: Total strikes across the country (lag 1)
- Contiguous regions: Strike counts in immediately adjacent areas
- Distance-weighted: Inverse distance-weighted strikes within 500km radius
- Capture spatial diffusion and regional coordination

**Geographic controls:**

- Fixed effects for ADM1 regions and countries
- Accounts for persistent regional heterogeneity

## Tidymodels overview

**Tidymodels** provides a unified framework for modeling in R:

- **recipes**: Feature engineering and preprocessing
- **parsnip**: Unified model specification interface
- **workflows**: Combine recipes and models
- **tune**: Hyperparameter optimization
- **yardstick**: Model evaluation metrics

Enables consistent workflow across different model types and engines.

## Tidymodels: Recipe example {.smaller}

```{r}
#| label: recipe-example
#| echo: true
#| eval: false

tree_recipe <- recipe(strike_count ~ ., data = pretest_df) |>
  update_role(date, new_role = "id") |>
  step_novel(action_geo_adm1_code, action_geo_country_code) |>
  step_other(action_geo_adm1_code, threshold = 0.02) |>
  step_mutate(is_year_end = as.numeric(is_year_end)) |>
  step_dummy(all_nominal_predictors()) |>
  step_impute_median(all_numeric_predictors()) |>
  step_lincomb(all_numeric_predictors()) |>
  step_corr(all_numeric_predictors(), threshold = 0.999) |>
  step_zv(all_predictors())
```

**Key steps**: Handle novel levels, pool rare categories, create dummies, impute missing values, remove collinearity

## Tidymodels: Model specification {.smaller}

```{r}
#| label: model-spec-example
#| echo: true
#| eval: false

# Linear model on log-transformed target
lm_log_spec <- linear_reg() |> 
  set_engine("lm") |> 
  set_mode("regression")

# LightGBM with Tweedie objective for count data
lgbm_twd_spec <- boost_tree(
  trees = tune(), 
  tree_depth = tune(), 
  min_n = tune(),
  loss_reduction = tune(), 
  sample_size = tune(), 
  mtry = tune(), 
  learn_rate = tune()
) |>
  set_engine("lightgbm", 
             objective = "tweedie", 
             tweedie_variance_power = 1.3,
             num_threads = 1)
```

**Parsnip** provides unified interface across engines (lm, glmnet, xgboost, lightgbm, etc.)

## Tidymodels: Workflows {.smaller}

```{r}
#| label: workflow-example
#| echo: true
#| eval: false

# Combine model specification and recipe
wf_lm_log <- workflow() |> 
  add_model(lm_log_spec) |> 
  add_recipe(lm_log_recipe)

# Tune hyperparameters on resamples
tuned_lgbm <- tune_grid(
  wf_lgbm_twd,
  resamples = val_resamples,
  grid = grid_boosted,
  metrics = metric_set(rmse, mae, rsq),
  control = control_grid(save_pred = FALSE, verbose = TRUE)
)

# Select best configuration
best_config <- select_best(tuned_lgbm, metric = "rmse")
final_wf <- finalize_workflow(wf_lgbm_twd, best_config)
```

**Workflows** bundle preprocessing and modeling for reproducible pipelines

## Cross-validation strategy {.smaller}

**Expanding window approach:**

- 3 rolling-origin folds
- Each fold: 26-week assessment period
- Training set grows with each fold (preserves temporal ordering)
- Uses all available historical data for each forecast

**Why expanding?**

- Respects real-world time series structure
- Maximizes training data for recent predictions
- Avoids look-ahead bias

## Hyperparameter tuning {.smaller}

**Boosted trees (LightGBM/XGBoost):** 24 configurations

- `trees`: 200-600
- `tree_depth`: 3-8
- `min_n`: 2-20
- `loss_reduction`: 0-5
- `sample_size`: 0.6-0.9
- `mtry`: 2-12
- `learn_rate`: 0.001-0.1 (log scale)

**GLMNET:** 50 configurations (10 penalty × 5 mixture levels)

**Selection metrics:** Both RMSE and MAE tested (results robust)

## Log model bias correction {.smaller}

**Problem:** log1p transformation introduces bias on back-transform

- Naive: ŷ = exp(log_pred) - 1
- Jensen's inequality: E[exp(X)] ≠ exp(E[X])

**Solution:** Duan's smearing estimator

1. Fit model on log1p(count)
2. Compute residuals: ε = log1p(count) - log_pred
3. Smearing factor: σ = mean(exp(ε))
4. Corrected prediction: ŷ = (exp(log_pred) × σ) - 1

**Result:** Unbiased predictions on original scale

# Part 4: Results

```{r}
#| label: visualization-setup

library(tidyverse)
library(ggplot2)
library(arrow)
library(patchwork)
library(scales)
library(viridis)

# Set theme
theme_set(theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  ))

# Define colors
model_colors <- c(
  "Single-stage" = viridis::turbo(10)[3],
  "Two-stage" = viridis::turbo(10)[8],
  "Baseline" = viridis::turbo(10)[2]
)

# Load data
sst_val_summary <- read_csv(file.path(root, "outputs/models/sst_val_summary_rmse.csv"), show_col_types = FALSE)
tsr_joint_opt <- read_csv(file.path(root, "outputs/models/tsr_joint_optimization_results__rmse.csv"), show_col_types = FALSE)
sst_naive_val <- read_csv(file.path(root, "outputs/models/sst_naive_baselines_validation.csv"), show_col_types = FALSE)

# Prepare single-stage data
sst_val_plot <- sst_val_summary |>
  mutate(
    model_type = "Single-stage",
    model_name = case_when(
      model == "lm_log" ~ "LM log1p",
      model == "glmnet_log" ~ "GLMNET log1p", 
      model == "lgbm_log" ~ "LightGBM log1p",
      model == "xgb_log" ~ "XGBoost log1p",
      model == "xgb_poi" ~ "XGBoost Poisson",
      model == "lgbm_poi" ~ "LightGBM Poisson",
      model == "lgbm_twd" ~ "LightGBM Tweedie",
      TRUE ~ model
    ),
    is_best = model == "lgbm_twd"
  )

# Prepare two-stage data
tsr_val_plot <- tsr_joint_opt |>
  head(10) |>
  mutate(
    model_type = "Two-stage",
    model_name = case_when(
      classifier == "glm" & severity_model == "gamma_xgb" ~ "GLM + Gamma XGBoost",
      classifier == "glm" & severity_model == "lgbm_log" ~ "GLM + LightGBM log1p",
      classifier == "glm" & severity_model == "gamma_lgbm" ~ "GLM + Gamma LightGBM",
      classifier == "glm" & severity_model == "xgb_log" ~ "GLM + XGBoost log1p",
      classifier == "glm" & severity_model == "glmnet_log" ~ "GLM + GLMNET log1p",
      classifier == "glm" & severity_model == "lm_log" ~ "GLM + LM log1p",
      classifier == "lgbm" & severity_model == "lm_log" ~ "LightGBM + LM log1p",
      classifier == "lgbm" & severity_model == "glmnet_log" ~ "LightGBM + GLMNET log1p",
      classifier == "lgbm" & severity_model == "gamma_lgbm" ~ "LightGBM + Gamma LightGBM",
      classifier == "lgbm" & severity_model == "gamma_xgb" ~ "LightGBM + Gamma XGBoost",
      TRUE ~ paste(classifier, "+", severity_model)
    ),
    is_best = row_number() == 1
  )
```

## Naive models

```{r}
#| label: fig0-naive-baselines
#| fig-align: center
#| fig-width: 10
#| fig-height: 6

# Prepare naive models data
naive_comparison <- sst_naive_val |>
  mutate(
    model_name = case_when(
      model == "naive_rolling_4wk" ~ "Rolling 4-week",
      model == "naive_rolling_8wk" ~ "Rolling 8-week",
      model == "naive_mean" ~ "Mean",
      model == "naive_median" ~ "Median",
      model == "naive_zero" ~ "Zero",
      model == "naive_persistence" ~ "Persistence",
      TRUE ~ model
    ),
    model_name = fct_reorder(model_name, -rmse)
  )

# Create plots
p1 <- ggplot(naive_comparison, aes(x = rmse, y = model_name)) +
  geom_point(size = 3, color = model_colors["Baseline"]) +
  labs(title = "RMSE") +
  theme(axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

p2 <- ggplot(naive_comparison, aes(x = mae, y = model_name)) +
  geom_point(size = 3, color = model_colors["Baseline"]) +
  labs(title = "MAE") +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5))

(p1 | p2)
```

## Validation results

```{r}
#| label: fig1-validation-all-models
#| fig-align: center

# Prepare baselines
baseline_val <- sst_naive_val |>
  filter(model %in% c("naive_rolling_8wk", "naive_mean")) |>
  mutate(
    model_type = "Baseline",
    model_name = case_when(
      model == "naive_rolling_8wk" ~ "Rolling 8wk",
      model == "naive_mean" ~ "Mean",
      TRUE ~ model
    ),
    is_best = model == "naive_rolling_8wk"
  )

# Combine all
val_comparison <- bind_rows(
  sst_val_plot |> select(model_name, model_type, rmse, mae, rsq, is_best),
  tsr_val_plot |> select(model_name, model_type, rmse = combined_rmse, mae = combined_mae, rsq = combined_rsq, is_best),
  baseline_val |> select(model_name, model_type, rmse, mae, rsq, is_best)
) |>
  mutate(
    model_name = fct_reorder(model_name, -rmse),
    model_type = fct_relevel(model_type, "Single-stage", "Two-stage", "Baseline"),
    selection_status = if_else(is_best, "Selected Model", "Not Selected")
  )

# Create plots
p1 <- ggplot(val_comparison, aes(x = rmse, y = model_name, color = model_type, size = selection_status)) +
  geom_point() +
  scale_color_manual(values = model_colors, name = "Model Type") +
  scale_size_manual(values = c("Not Selected" = 2, "Selected Model" = 3), name = "Selection") +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2)) +
  labs(title = "RMSE") +
  theme(legend.position = "right",
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

p2 <- ggplot(val_comparison, aes(x = mae, y = model_name, color = model_type, size = selection_status)) +
  geom_point() +
  scale_color_manual(values = model_colors, name = "Model Type") +
  scale_size_manual(values = c("Not Selected" = 2, "Selected Model" = 3), name = "Selection") +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2)) +
  labs(title = "MAE") +
  theme(legend.position = "right",
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5))

p3 <- ggplot(val_comparison, aes(x = rsq, y = model_name, color = model_type, size = selection_status)) +
  geom_point() +
  scale_color_manual(values = model_colors, name = "Model Type") +
  scale_size_manual(values = c("Not Selected" = 2, "Selected Model" = 3), name = "Selection") +
  scale_x_continuous(labels = percent_format()) +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2)) +
  labs(title = "R²") +
  theme(legend.position = "right",
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5))

(p1 | p2 | p3) + plot_layout(guides = "collect") & theme(legend.position = "right")
```

## Non-zero performance comparison

```{r}
#| label: fig3-nonzero-performance
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-width: 10
#| fig-height: 6

# Load data
sst_val_separate <- read_csv(file.path(root, "outputs/models/sst_val_separate_metrics_rmse.csv"), show_col_types = FALSE)
tsr_stage2_metrics <- read_csv(file.path(root, "outputs/models/tsr_stage2_severity_metrics__rmse.csv"), show_col_types = FALSE)

# All models list
all_models <- c(
  "LM log1p", "GLMNET log1p", "LightGBM log1p", "XGBoost log1p",
  "XGBoost Poisson", "LightGBM Tweedie",
  "Gamma LightGBM", "Gamma XGBoost"
)

# Prepare single-stage data
sst_nonzero_plot <- sst_val_separate |>
  filter(week_type == "non_zero", .metric == "rmse", model != "lgbm_poi") |>
  select(model, rmse = .estimate) |>
  mutate(
    model_name = case_when(
      model == "lm_log" ~ "LM log1p",
      model == "glmnet_log" ~ "GLMNET log1p", 
      model == "lgbm_log" ~ "LightGBM log1p",
      model == "xgb_log" ~ "XGBoost log1p",
      model == "xgb_poi" ~ "XGBoost Poisson",
      model == "lgbm_twd" ~ "LightGBM Tweedie",
      TRUE ~ model
    ),
    model_type = "Single-stage",
    is_best = model == "lgbm_twd"
  )

# Prepare two-stage data
tsr_stage2_plot <- tsr_stage2_metrics |>
  select(model, rmse) |>
  mutate(
    model_name = case_when(
      model == "gamma_lgbm" ~ "Gamma LightGBM",
      model == "gamma_xgb" ~ "Gamma XGBoost", 
      model == "lgbm_log" ~ "LightGBM log1p",
      model == "xgb_log" ~ "XGBoost log1p",
      model == "glmnet_log" ~ "GLMNET log1p",
      model == "lm_log" ~ "LM log1p",
      TRUE ~ model
    ),
    model_type = "Stage 2",
    is_best = model == "gamma_lgbm"
  )

# Create comparison dataframe
comparison_data <- tibble(
  model_name = factor(all_models, levels = rev(all_models)),
  single_stage_rmse = NA_real_,
  stage2_rmse = NA_real_,
  single_stage_selected = FALSE,
  stage2_selected = FALSE
)

# Fill in values
for(i in seq_along(sst_nonzero_plot$model_name)) {
  idx <- which(comparison_data$model_name == sst_nonzero_plot$model_name[i])
  if(length(idx) > 0) {
    comparison_data$single_stage_rmse[idx] <- sst_nonzero_plot$rmse[i]
    comparison_data$single_stage_selected[idx] <- sst_nonzero_plot$is_best[i]
  }
}

for(i in seq_along(tsr_stage2_plot$model_name)) {
  idx <- which(comparison_data$model_name == tsr_stage2_plot$model_name[i])
  if(length(idx) > 0) {
    comparison_data$stage2_rmse[idx] <- tsr_stage2_plot$rmse[i]
    comparison_data$stage2_selected[idx] <- tsr_stage2_plot$is_best[i]
  }
}

# Pivot to long
plot_data <- comparison_data |>
  pivot_longer(
    cols = c(single_stage_rmse, stage2_rmse),
    names_to = "model_type",
    values_to = "rmse",
    values_drop_na = TRUE
  ) |>
  mutate(
    model_type = case_when(
      model_type == "single_stage_rmse" ~ "Single-stage",
      model_type == "stage2_rmse" ~ "Stage 2"
    ),
    is_selected = case_when(
      model_type == "Single-stage" ~ single_stage_selected,
      model_type == "Stage 2" ~ stage2_selected
    ),
    selection_status = if_else(is_selected, "Selected Model", "Not Selected")
  )

# Create plot
ggplot(plot_data, aes(x = rmse, y = model_name, color = model_type, size = selection_status)) +
  geom_point() +
  scale_color_manual(values = c("Single-stage" = viridis::turbo(10)[3], "Stage 2" = viridis::turbo(10)[8]), name = "Model Type") +
  scale_size_manual(values = c("Not Selected" = 3.5, "Selected Model" = 5), name = "Selection") +
  guides(color = guide_legend(order = 1), size = guide_legend(order = 2)) +
  labs(
    x = "RMSE",
    y = "Model"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5)
  )
```

## Test set performance (best model)

```{r}
#| label: fig2-test-best-models
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

# Load test data
sst_model_metadata <- read_csv(file.path(root, "outputs/models/sst_model_metadata_rmse.csv"), show_col_types = FALSE)
tsr_test_combined <- read_csv(file.path(root, "outputs/models/two_stage_refined_test_combined__rmse.csv"), show_col_types = FALSE)
sst_naive_test <- read_csv(file.path(root, "outputs/models/sst_naive_baselines_test.csv"), show_col_types = FALSE)

# Prepare data
best_models <- tibble(
  model_name = c("LightGBM Tweedie", "GLM + Gamma LGBM", "Rolling 8wk"),
  model_type = c("Single-stage", "Two-stage", "Baseline"),
  val_rmse = c(sst_val_summary$rmse[sst_val_summary$model == "lgbm_twd"], 
               tsr_joint_opt$combined_rmse[1],
               sst_naive_val$rmse[sst_naive_val$model == "naive_rolling_8wk"]),
  test_rmse = c(sst_model_metadata$test_rmse[1],
                tsr_test_combined$.estimate[tsr_test_combined$.metric == "rmse"],
                sst_naive_test$rmse[sst_naive_test$model == "naive_rolling_8wk"]),
  val_mae = c(sst_val_summary$mae[sst_val_summary$model == "lgbm_twd"],
              tsr_joint_opt$combined_mae[1], 
              sst_naive_val$mae[sst_naive_val$model == "naive_rolling_8wk"]),
  test_mae = c(sst_model_metadata$test_mae[1],
               tsr_test_combined$.estimate[tsr_test_combined$.metric == "mae"],
               sst_naive_test$mae[sst_naive_test$model == "naive_rolling_8wk"]),
  val_rsq = c(sst_val_summary$rsq[sst_val_summary$model == "lgbm_twd"],
              tsr_joint_opt$combined_rsq[1],
              sst_naive_val$rsq[sst_naive_val$model == "naive_rolling_8wk"]),
  test_rsq = c(sst_model_metadata$test_rsq[1],
               tsr_test_combined$.estimate[tsr_test_combined$.metric == "rsq"],
               sst_naive_test$rsq[sst_naive_test$model == "naive_rolling_8wk"])
) |>
  pivot_longer(cols = starts_with(c("val_", "test_")), 
               names_to = c("dataset", "metric"), 
               names_sep = "_") |>
  pivot_wider(names_from = metric, values_from = value) |>
  mutate(
    model_name = fct_relevel(model_name, "LightGBM Tweedie", "GLM + Gamma LGBM", "Rolling 8wk"),
    dataset = fct_relevel(dataset, "val", "test")
  )

# Create plots
p_rmse <- ggplot(best_models, aes(x = model_name, y = rmse, fill = dataset)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("val" = viridis::turbo(10)[1], "test" = viridis::turbo(10)[9]), 
                    labels = c("Validation", "Test")) +
  labs(title = "RMSE", fill = "Dataset") +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

p_mae <- ggplot(best_models, aes(x = model_name, y = mae, fill = dataset)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("val" = viridis::turbo(10)[1], "test" = viridis::turbo(10)[9]),
                    labels = c("Validation", "Test")) +
  labs(title = "MAE", fill = "Dataset") +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

p_rsq <- ggplot(best_models, aes(x = model_name, y = rsq, fill = dataset)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("val" = viridis::turbo(10)[1], "test" = viridis::turbo(10)[9]),
                    labels = c("Validation", "Test")) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "R²", fill = "Dataset") +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

(p_rmse | p_mae | p_rsq) + plot_layout(guides = "collect")
```

## Predictions vs actuals (Tweedie)

```{r}
#| label: fig4-time-series
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-width: 18
#| fig-height: 7

# Load data
tsr_test_predictions <- read_parquet(file.path(root, "outputs/models/two_stage_refined_test_predictions__rmse.parquet"))
sst_predictions <- read_parquet(file.path(root, "outputs/models/sst_predictions__rmse.parquet"))

# Aggregate to weekly totals
weekly_tsr <- tsr_test_predictions |>
  group_by(date) |>
  summarise(
    actual_total = sum(actual_count, na.rm = TRUE),
    tsr_pred_total = sum(expected_count, na.rm = TRUE),
    .groups = "drop"
  )

weekly_sst <- sst_predictions |>
  group_by(date) |>
  summarise(
    sst_pred_total = sum(.pred, na.rm = TRUE),
    .groups = "drop"
  )

# Join
time_series_data <- weekly_tsr |>
  left_join(weekly_sst, by = "date") |>
  select(date, actual = actual_total, sst_pred = sst_pred_total) |>
  pivot_longer(cols = c(actual, sst_pred), 
               names_to = "series", values_to = "count") |>
  mutate(
    series = case_when(
      series == "actual" ~ "Actual",
      series == "sst_pred" ~ "Single-stage (LightGBM Tweedie)"
    ),
    series = fct_relevel(series, "Actual", "Single-stage (LightGBM Tweedie)")
  )

# Time series plot
fig4a <- ggplot(time_series_data, aes(x = date, y = count, color = series, linetype = series)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = c("Actual" = "black", 
                                "Single-stage (LightGBM Tweedie)" = viridis::turbo(10)[3])) +
  scale_linetype_manual(values = c("Actual" = "solid",
                                   "Single-stage (LightGBM Tweedie)" = "solid")) +
  labs(
    title = "A) Weekly Aggregated Time Series",
    subtitle = "Sum across all 1,550 ADM1 regions per week",
    x = "Date",
    y = "Total Strike Count",
    color = "Series",
    linetype = "Series"
  ) +
  theme(legend.position = "bottom")

# Region-level scatter
tsr_data <- tsr_test_predictions |>
  select(actual_count, expected_count)

sst_data <- sst_predictions |>
  select(.pred)

region_data <- bind_cols(
  tsr_data,
  sst_data |> rename(sst_pred = .pred)
) |>
  select(actual = actual_count, sst_pred)

# Scatter plot
fig4b <- ggplot(region_data, aes(x = actual, y = sst_pred)) +
  geom_hex(bins = 30, alpha = 0.8) +
  scale_fill_viridis_c(option = "turbo", name = "Count", trans = "log10",
                       begin = 0.2, end = 0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 1, color = "black") +
  labs(
    title = "B) Region-Level Calibration",
    subtitle = "80,600 region-week predictions",
    x = "Actual Strike Count",
    y = "Predicted Strike Count"
  ) +
  theme(legend.position = "right")

# Combine
library(patchwork)
fig4a | fig4b
```

## Feature importance (Tweedie)

```{r}
#| label: fig5-feature-importance
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

library(tidymodels)
library(vip)

# Load model
best_model_fit <- readRDS(file.path(root, "outputs/models/sst_best_fitted_model_rmse.rds"))

# Extract feature importance from LightGBM Tweedie model
importance_df <- vip::vi_model(best_model_fit) |>
  arrange(desc(Importance)) |>
  head(15) |>
  mutate(
    feature_clean = case_when(
      str_detect(Variable, "strike_count_lag1") ~ "Strike Count (lag 1)",
      str_detect(Variable, "strike_count_lag2") ~ "Strike Count (lag 2)", 
      str_detect(Variable, "strike_count_lag4") ~ "Strike Count (lag 4)",
      str_detect(Variable, "rolling_avg_4wk") ~ "Rolling 4-week Average",
      str_detect(Variable, "rolling_avg_8wk") ~ "Rolling 8-week Average",
      str_detect(Variable, "total_articles_lag1") ~ "Total Articles (lag 1)",
      str_detect(Variable, "total_articles_lag2") ~ "Total Articles (lag 2)",
      str_detect(Variable, "total_mentions_lag1") ~ "Total Mentions (lag 1)",
      str_detect(Variable, "total_mentions_lag2") ~ "Total Mentions (lag 2)",
      str_detect(Variable, "avg_tone_lag1") ~ "Average Tone (lag 1)",
      str_detect(Variable, "avg_tone_lag2") ~ "Average Tone (lag 2)",
      str_detect(Variable, "actor_diversity_lag1") ~ "Actor Diversity (lag 1)",
      str_detect(Variable, "country_strikes_lag1") ~ "Country Strikes (lag 1)",
      str_detect(Variable, "contig_strikes_t1") ~ "Contiguous Strikes (t-1)",
      str_detect(Variable, "distw_strikes_t1") ~ "Distance-weighted Strikes (t-1)",
      str_detect(Variable, "time_trend") ~ "Time Trend",
      TRUE ~ str_replace_all(Variable, "_", " ") |> str_to_title()
    ),
    feature_clean = fct_reorder(feature_clean, Importance)
  )

# Create plot
ggplot(importance_df, aes(x = Importance, y = feature_clean)) +
  geom_col(alpha = 0.8, fill = viridis::turbo(10)[3]) +
  labs(
    title = "Feature Importance: LightGBM Tweedie",
    subtitle = "Gain-based importance for top 15 predictors",
    x = "Importance (Gain)",
    y = "Feature"
  ) +
  theme(
    legend.position = "right",
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

## Key findings {.smaller}

**Model performance:**

- Best model: **LightGBM Tweedie** (test RMSE: 0.257, 2% better than baseline)
- Two-stage models showed promise on validation but collapsed on test data
- Single-stage Tweedie proved more robust to distributional shift

**Distributional shift challenge:**

- Training data (2015-2024): 2.2% strike rate, mean count = 0.037
- Test data (late 2024-2025): 1.5% strike rate, mean count = 0.023
- **33% decline in strike frequency** between training and test periods
- Two-stage models failed to adapt, predicting near-constant values

## Key findings {.smaller}

**Important predictors:**

- Temporal lags (lag1, rolling averages) dominate importance
- Media coverage (articles, mentions) moderately important  
- Spatial spillovers (contiguous, distance-weighted) contribute
- Geographic fixed effects capture regional heterogeneity

**Takeaway:** Single-stage models may handle rare events more robustly than hurdle approaches when base rates are extremely low and unstable

# Part 5: Future Directions

## Improve data

- **LLM-based filtering**: Use GPT/Claude to classify GKG themes and article content for strike relevance
- **Article metadata analysis**: Extract structured data (industries, actors, strike types) from news articles  
- **Multi-stage validation**: Combine GDELT codes + GKG themes + LLM classification for higher precision
- **Active learning**: Use LLM confidence scores to identify borderline cases for human review

## Add external features

- **V-Dem**: Democracy indices, electoral competitiveness, civil liberties
- **World Bank**: Unemployment, inflation, GDP growth, labor market indicators
- **ILO**: Union density, collective bargaining coverage, labor regulations
- **ACLED**: Protest intensity, government repression

## Add external features

**Challenges**: 

- No direct merge key (GDELT uses bespoke ADM1 codes)
- Temporal alignment (quarterly/annual vs weekly data)
- Missing data

## Narrow geographic focus

<br>

- Train only on regions with >1 strike per year (reduces dataset but improves signal)
- Use country-level strike intensity as screening criteria
- Focus on historically active labor regions (industrial centers, capitals)

## GDELT tone measurement

<br>

**Current**: Dictionary-based sentiment (-100 to +100 scale, typically -10 to +10)

**Limitations**: No context awareness, static dictionary

**Modern ML**: Transformer models + LLMs with few-shot prompting for better context and domain adaptation

## Capturing strike bursts

<br>

- **Quantile regression**: Predict 75th percentile instead of mean
- **Momentum features**: Strike acceleration, rolling maximums  
- **Extreme event classifier**: Separate model for counts > 5

## Capturing strike bursts

<br> 

- Higher-threshold hurdle models (split at count > 2)
- Asymmetric loss functions (penalize underprediction more heavily)
- Ensemble quantile predictions (multiple percentiles)
