---
title: "Two-Stage Strike Forecasting Analysis"
format: html
---

## Setup

```{r}
#| label: setup
#| message: false

library(tidymodels)
library(tidyverse)
library(lubridate)
library(arrow)
library(bonsai)
library(finetune)
library(themis)
library(janitor)
library(doFuture)   # Parallel backend for tune_grid via foreach/future
library(vip)

tidymodels_prefer()
conflicted::conflicts_prefer(dplyr::lag)
conflicted::conflicts_prefer(stringr::fixed)
set.seed(2024)

# Parallel tuning setup (Apple Silicon safe)
# We parallelize the OUTER loop (resamples × grid) and restrict per-model threads to avoid oversubscription.
# Adjust workers if you see high RAM/CPU: 2–4 is a good range on M1.
registerDoFuture()
plan(multisession, workers = 4)  # 3–4 workers often OK on Apple Silicon; reduce if RAM pressure

# Allow larger globals transfer to workers (resample objects can be large)
options(future.globals.maxSize = 8 * 1024^3)  # 8 GiB

# Racing control is created after resamples are defined so burn-in can adapt
ctrl_race <- NULL
```

## Load data

```{r}
#| label: load-data

adm_full <- read_parquet("../../data/analysis/adm_week_full.parquet") |>
  clean_names()

glimpse(adm_full)
```

## Targets and predictors

```{r}
#| label: targets-predictors

# Occurrence target
adm_full <- adm_full |>
  mutate(y_occ = factor(if_else(strike_count > 0, "yes", "no"), levels = c("no", "yes")))

# Predictor set: lag-only + time + spatial (avoid contemporaneous leakage)
predictor_cols <- adm_full |>
  select(
    # IDs
    action_geo_adm1_code, action_geo_country_code,
    # time
    time_trend, year_sin, year_cos, month, quarter, is_year_end,
    # lagged counts & rolling
    strike_count_lag1, strike_count_lag2, strike_count_lag4,
    rolling_avg_4wk, rolling_avg_8wk,
    # media/tone lags
    total_articles_lag1, total_articles_lag2,
    total_mentions_lag1, total_mentions_lag2,
    avg_tone_lag1, avg_tone_lag2,
    # actor lags
    actor_diversity_lag1,
    unique_actor1_types_lag1, unique_actor2_types_lag1,
    prop_gov_lag1, prop_labor_lag1, prop_civil_lag1,
    # spatial lags
    country_strikes_lag1, contig_strikes_t1, distw_strikes_t1
  ) |>
  names()

model_df <- adm_full |>
  select(date, strike_count, y_occ, all_of(predictor_cols)) |>
  arrange(date)

# Severity subset (positive weeks only), after feature construction
severity_df <- model_df |>
  filter(strike_count > 0)
```

## Time splits: rolling-origin CV + final 52-week holdout

```{r}
#| label: splits

all_weeks <- sort(unique(model_df$date))
test_weeks <- tail(all_weeks, 52)

pretest_df <- model_df |>
  filter(date < min(test_weeks))
test_df <- model_df |>
  filter(date %in% test_weeks)

# Build rolling-origin splits on pretest period (3 folds × 26-week assessments)
assessment_weeks <- 26
n_folds <- 3

pretest_max <- max(pretest_df$date)
assess_end_dates <- pretest_max - weeks(rev(seq(0, by = assessment_weeks, length.out = n_folds)))
assess_start_dates <- assess_end_dates - weeks(assessment_weeks - 1)

cap_weeks <- 156  # cap analysis window to last N weeks (~3 years)

make_time_split <- function(df, start_date, end_date) {
  window_start <- start_date - weeks(cap_weeks)
  analysis_idx <- which(df$date >= window_start & df$date < start_date)
  assessment_idx <- which(df$date >= start_date & df$date <= end_date)
  rsample::make_splits(list(analysis = analysis_idx, assessment = assessment_idx), df)
}

val_splits <- map2(assess_start_dates, assess_end_dates, ~ make_time_split(pretest_df, .x, .y))
val_resamples <- rsample::manual_rset(val_splits, paste0("fold", seq_along(val_splits)))

cat("Validation folds:", nrow(val_resamples), "(assessment window", assessment_weeks, "weeks)\n")
```

## Recipes

```{r}
#| label: recipes

# Common preprocessor (dummy categorical, numeric median impute, encode logical)
base_recipe <- recipe(~ ., data = pretest_df) |>
  update_role(date, new_role = "id") |>
  step_novel(action_geo_adm1_code, action_geo_country_code) |>
  step_other(action_geo_adm1_code, threshold = 0.02) |>
  step_dummy(action_geo_adm1_code, action_geo_country_code) |>
  step_mutate(is_year_end = as.numeric(is_year_end)) |>
  step_dummy(all_nominal_predictors()) |>
  step_impute_median(all_numeric_predictors()) |>
  step_zv(all_predictors())

# Occurrence recipes (predict y_occ)
# - Downsampled for Logistic/RF
# - Weighted (no downsample) for XGB/LGBM
occ_recipe_down <- base_recipe |>
  update_role(y_occ, new_role = "outcome") |>
  step_rm(strike_count) |>
  step_downsample(y_occ, under_ratio = 5)

occ_recipe_wt <- base_recipe |>
  update_role(y_occ, new_role = "outcome") |>
  step_rm(strike_count)

# Severity recipes: define explicitly so `y_occ` is removed BEFORE dummying
sev_recipe_log <- recipe(strike_count ~ ., data = pretest_df) |>
  update_role(date, new_role = "id") |>
  step_rm(y_occ) |>
  step_novel(action_geo_adm1_code, action_geo_country_code) |>
  step_other(action_geo_adm1_code, threshold = 0.02) |>
  step_dummy(action_geo_adm1_code, action_geo_country_code) |>
  step_mutate(is_year_end = as.numeric(is_year_end)) |>
  step_dummy(all_nominal_predictors()) |>
  step_impute_median(all_numeric_predictors()) |>
  step_zv(all_predictors()) |>
  step_log(strike_count, offset = 1, skip = TRUE)

sev_recipe_tree <- recipe(strike_count ~ ., data = pretest_df) |>
  update_role(date, new_role = "id") |>
  step_rm(y_occ) |>
  step_novel(action_geo_adm1_code, action_geo_country_code) |>
  step_other(action_geo_adm1_code, threshold = 0.02) |>
  step_dummy(action_geo_adm1_code, action_geo_country_code) |>
  step_mutate(is_year_end = as.numeric(is_year_end)) |>
  step_dummy(all_nominal_predictors()) |>
  step_impute_median(all_numeric_predictors()) |>
  step_zv(all_predictors())
```

## Model specifications

```{r}
#| label: model-specs

# Stage 1: Occurrence (classification)
# Compute imbalance weight for XGBoost on pretest data
pos_count <- sum(pretest_df$y_occ == "yes")
neg_count <- sum(pretest_df$y_occ == "no")
scale_pos_wt <- ifelse(pos_count > 0, neg_count / pos_count, 1)
log_reg_spec <- logistic_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet") |>
  set_mode("classification")

rf_clf_spec <- rand_forest(trees = 500, mtry = tune(), min_n = tune()) |>
  set_engine("ranger", classification = TRUE, num.threads = 1) |>
  set_mode("classification")

xgb_clf_spec <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_engine("xgboost", objective = "binary:logistic", tree_method = "hist", nthread = 1, scale_pos_weight = !!scale_pos_wt) |>
  set_mode("classification")

lgb_clf_spec <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_engine("lightgbm", objective = "binary", metric = "aucpr", num_threads = 1, is_unbalance = TRUE) |>
  set_mode("classification")

# Stage 2: Severity (regression)
lm_log_spec <- linear_reg() |>
  set_engine("lm") |>
  set_mode("regression")

xgb_poi_spec <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_engine("xgboost", objective = "count:poisson", tree_method = "hist", nthread = 1) |>
  set_mode("regression")

lgb_poi_spec <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_engine("lightgbm", objective = "poisson", num_threads = 1) |>
  set_mode("regression")

lgb_twd_spec <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_engine("lightgbm", objective = "tweedie", tweedie_variance_power = 1.5, num_threads = 1) |>
  set_mode("regression")

# Grids
# Separate grids so each model only receives parameters it supports
grid_boosted <- grid_space_filling(
  trees(range = c(200, 600)),
  tree_depth(range = c(3, 8)),
  min_n(range = c(2, 20)),
  loss_reduction(range = c(0, 5)),
  sample_size = sample_prop(range = c(0.6, 0.9)),
  mtry(range = c(2, min(12, length(predictor_cols)))),
  learn_rate(range = c(-3, -1), trans = log10_trans()),
  size = 12
)

grid_rf <- grid_space_filling(
  mtry(range = c(5, 12)),
  min_n(range = c(2, 20)),
  size = 12
)

lambda_grid <- grid_regular(penalty(range = c(-5, 0)), levels = 10)
```

## Workflows

```{r}
#| label: workflows

occ_workflows <- list(
  log_reg = workflow() |> add_model(log_reg_spec) |> add_recipe(occ_recipe_down),
  rf = workflow() |> add_model(rf_clf_spec) |> add_recipe(occ_recipe_down),
  xgb = workflow() |> add_model(xgb_clf_spec) |> add_recipe(occ_recipe_wt),
  lgb = workflow() |> add_model(lgb_clf_spec) |> add_recipe(occ_recipe_wt)
)

sev_workflows <- list(
  lm_log = workflow() |> add_model(lm_log_spec) |> add_recipe(sev_recipe_log),
  xgb_poi = workflow() |> add_model(xgb_poi_spec) |> add_recipe(sev_recipe_tree),
  lgb_poi = workflow() |> add_model(lgb_poi_spec) |> add_recipe(sev_recipe_tree),
  lgb_twd = workflow() |> add_model(lgb_twd_spec) |> add_recipe(sev_recipe_tree)
)
```

## Tuning and evaluation on rolling folds

```{r}
#| label: tuning-eval
#| cache: false

# Metrics (use F2 for recall emphasis)
f_meas2 <- yardstick::metric_tweak("f_meas2", yardstick::f_meas, beta = 2)
# Force positive class to be "yes" (second level) for AUC metrics
pr_auc_yes <- yardstick::metric_tweak("pr_auc", yardstick::pr_auc, event_level = "second")
roc_auc_yes <- yardstick::metric_tweak("roc_auc", yardstick::roc_auc, event_level = "second")
# Keep combined metrics for tuning (AUCs computed for positive class = "yes")
clf_metrics <- metric_set(pr_auc_yes, roc_auc_yes, f_meas2)
# Separate prob- and class-based sets for fold evaluation to avoid mixed mappings
clf_metrics_prob <- metric_set(pr_auc_yes, roc_auc_yes)
clf_metrics_class <- metric_set(f_meas2)
# Put RMSE first so the race and ranking metric agree
reg_metrics <- metric_set(rmse, mae, rsq)

# Tune classifiers
# Create racing control after resamples so burn-in < nrow(val_resamples)
ctrl_race <- control_race(
  save_pred = FALSE,
  verbose = TRUE,
  allow_par = TRUE,
  parallel_over = "resamples",
  save_workflow = FALSE,
  burn_in = max(1, min(2, nrow(val_resamples) - 1))
)
tuned_clf <- imap(occ_workflows, function(wf, name) {
  cat("[Stage 1] Tuning classifier:", name, "\n")
  res <- if (name == "log_reg") {
    tune_race_anova(wf, resamples = val_resamples, grid = lambda_grid, metrics = clf_metrics,
                    control = ctrl_race)
  } else if (name == "rf") {
    tune_race_anova(wf, resamples = val_resamples, grid = grid_rf, metrics = clf_metrics,
                    control = ctrl_race)
  } else {
    tune_race_anova(wf, resamples = val_resamples, grid = grid_boosted, metrics = clf_metrics,
                    control = ctrl_race)
  }
  cat("[Stage 1] Completed classifier:", name, "\n")
  res
})

# Inspect tuning notes for failures/warnings (optional)
try({
  show_notes(.Last.tune.result)
}, silent = TRUE)

# Tune regressors (on positive weeks only within each fold)
# Build positive-only resamples mirroring val_resamples by dates
pos_pretest_df <- pretest_df |>
  filter(strike_count > 0)

map_split_to_pos <- function(split) {
  analysis_dates <- rsample::analysis(split)$date
  assessment_dates <- rsample::assessment(split)$date
  pos_analysis_idx <- which(pos_pretest_df$date %in% analysis_dates)
  pos_assessment_idx <- which(pos_pretest_df$date %in% assessment_dates)
  rsample::make_splits(list(analysis = pos_analysis_idx, assessment = pos_assessment_idx), pos_pretest_df)
}

pos_val_splits <- purrr::map(val_resamples$splits, map_split_to_pos)
pos_val_resamples <- rsample::manual_rset(pos_val_splits, val_resamples$id)

tuned_reg <- imap(sev_workflows, function(wf, name) {
  if (name == "lm_log") {
    cat("[Stage 2] Skipping tuning for:", name, "(no hyperparameters)\n")
    list(workflow = wf)
  } else {
    cat("[Stage 2] Tuning regressor:", name, "\n")
    res <- tune_race_anova(wf, resamples = pos_val_resamples, grid = grid_boosted, metrics = reg_metrics,
                           control = ctrl_race)
    cat("[Stage 2] Completed regressor:", name, "\n")
    res
  }
})

# Select best per model
best_clf <- imap(tuned_clf, ~ select_best(.x, metric = "pr_auc"))
cat("[Stage 1] Selected best hyperparameters for all classifiers (metric=pr_auc).\n")

# Use workflow names from imap (.y) to index best_clf, not cur_group_id()
finalized_clf <- imap(occ_workflows, ~ finalize_workflow(.x, best_clf[[.y]]))

finalized_reg <- list(
  lm_log = sev_workflows$lm_log,
  xgb_poi = finalize_workflow(sev_workflows$xgb_poi, select_best(tuned_reg$xgb_poi, metric = "rmse")),
  lgb_poi = finalize_workflow(sev_workflows$lgb_poi, select_best(tuned_reg$lgb_poi, metric = "rmse")),
  lgb_twd = finalize_workflow(sev_workflows$lgb_twd, select_best(tuned_reg$lgb_twd, metric = "rmse"))
)

# Helper to predict severity for positive weeks and back-transform if needed
predict_severity <- function(fit, new_data, model_name) {
  if (model_name == "lm_log") {
    pred_log <- predict(fit, new_data = new_data)$.pred
    pred <- pmax(expm1(pred_log), 0)
    pred[is.na(pred)] <- 0
    pred
  } else {
    pred <- predict(fit, new_data = new_data)$.pred
    pred[is.na(pred)] <- 0
    pred
  }
}

# Print per-fold metrics for selected configs before moving on
cat("\n[Stage 1] Per-fold metrics for best classifiers:\n")
clf_fold_metrics <- purrr::imap_dfr(tuned_clf, function(res, name) {
  best_cfg <- best_clf[[name]]$.config
  tune::collect_metrics(res, summarize = FALSE) |>
    dplyr::filter(.config == best_cfg, .metric %in% c("pr_auc", "roc_auc", "f_meas2")) |>
    dplyr::select(id, .metric, .estimate) |>
    tidyr::pivot_wider(names_from = .metric, values_from = .estimate) |>
    dplyr::mutate(model = name, .before = 1)
}) |>
  dplyr::arrange(model, id)
print(clf_fold_metrics)

cat("\n[Stage 2] Per-fold metrics for best regressors (positives only):\n")
reg_names_tuned <- setdiff(names(tuned_reg), "lm_log")
reg_fold_metrics <- purrr::imap_dfr(tuned_reg[reg_names_tuned], function(res, name) {
  best_cfg <- dplyr::pull(select_best(res, metric = "rmse"), .config)
  tune::collect_metrics(res, summarize = FALSE) |>
    dplyr::filter(.config == best_cfg, .metric %in% c("rmse", "mae", "rsq")) |>
    dplyr::select(id, .metric, .estimate) |>
    tidyr::pivot_wider(names_from = .metric, values_from = .estimate) |>
    dplyr::mutate(model = name, .before = 1)
}) |>
  dplyr::arrange(model, id)
  
print(reg_fold_metrics)
```

## Evaluate all pipelines on validation folds

```{r}
# Evaluate 16 pipelines on validation folds
pipeline_results <- list()

for (clf_name in names(finalized_clf)) {
  for (reg_name in names(finalized_reg)) {
    cat("[Pipelines] Evaluating:", clf_name, "+", reg_name, "\n")
    # Fit on analysis portions and evaluate on assessment, per fold
    fold_metrics <- map2_dfr(val_resamples$splits, pos_val_resamples$splits, function(split_all, split_pos) {
      analysis_all <- analysis(split_all)
      assess_all <- assessment(split_all)
      analysis_pos <- analysis(split_pos)
      assess_pos <- assessment(split_pos)

      # Fit classifier on all weeks
      clf_fit <- fit(finalized_clf[[clf_name]], data = analysis_all)
      prob_yes <- predict(clf_fit, new_data = assess_all, type = "prob")$.pred_yes
      prob_yes[is.na(prob_yes)] <- 0
      prob_yes <- pmin(pmax(prob_yes, 0), 1)
      occ_pred <- if_else(prob_yes >= 0.5, "yes", "no") |> factor(levels = c("no", "yes"))
      # Compute classification metrics directly on vectors (more robust)
      truth_occ <- factor(assess_all$y_occ, levels = c("no", "yes"))
      pr_auc_val <- yardstick::pr_auc_vec(truth_occ, prob_yes, event_level = "second")
      roc_auc_val <- yardstick::roc_auc_vec(truth_occ, prob_yes, event_level = "second")
      f2_val <- yardstick::f_meas_vec(truth_occ, occ_pred, beta = 2)

      # Fit regressor on positive weeks
      reg_fit <- fit(finalized_reg[[reg_name]], data = analysis_pos)
      sev_hat_assess_pos <- predict_severity(reg_fit, new_data = assess_pos, model_name = reg_name)
      sev_fold <- reg_metrics(bind_cols(truth = assess_pos$strike_count, .pred = sev_hat_assess_pos),
                              truth = truth, estimate = .pred)

      # Combined expected counts on all assessment rows
      # Need severity predictions for all rows; use positive-only model and multiply by prob
      # For rows with zero true strikes, we still use expected count = p * E[Y|>0]
      sev_hat_all <- predict_severity(reg_fit, new_data = assess_all, model_name = reg_name)
      sev_hat_all <- pmax(pmin(sev_hat_all, quantile(sev_hat_all, 0.999, na.rm = TRUE)), 0)
      sev_hat_all[!is.finite(sev_hat_all)] <- 0
      combined_hat <- prob_yes * sev_hat_all
      combined_fold <- reg_metrics(bind_cols(truth = assess_all$strike_count, .pred = combined_hat),
                                   truth = truth, estimate = .pred)

      tibble(
        clf = clf_name,
        reg = reg_name,
        pr_auc = pr_auc_val,
        roc_auc = roc_auc_val,
        f2 = f2_val,
        sev_mae = sev_fold$.estimate[sev_fold$.metric == "mae"],
        sev_rmse = sev_fold$.estimate[sev_fold$.metric == "rmse"],
        sev_r2 = sev_fold$.estimate[sev_fold$.metric == "rsq"],
        comb_mae = combined_fold$.estimate[combined_fold$.metric == "mae"],
        comb_rmse = combined_fold$.estimate[combined_fold$.metric == "rmse"],
        comb_r2 = combined_fold$.estimate[combined_fold$.metric == "rsq"]
      )
    })

    pipeline_results[[paste(clf_name, reg_name, sep = "_")]] <- fold_metrics
    cat("[Pipelines] Completed:", clf_name, "+", reg_name, "\n")
  }
}

val_summary <- bind_rows(pipeline_results, .id = "pipeline") |>
  group_by(pipeline, clf, reg) |>
  summarise(across(where(is.numeric), mean, na.rm = TRUE), .groups = "drop") |>
  arrange(comb_rmse)

print(val_summary)

# Tune decision thresholds for classifiers on validation folds (optimize F2)
threshold_grid <- seq(0.05, 0.95, by = 0.05)

compute_best_threshold <- function(wf) {
  preds_tbl <- map_dfr(val_resamples$splits, function(split) {
    analysis_all <- analysis(split)
    assess_all <- assessment(split)
    fit_obj <- fit(wf, data = analysis_all)
    tibble(
      truth = assess_all$y_occ,
      prob = predict(fit_obj, new_data = assess_all, type = "prob")$.pred_yes
    )
  })

  f_by_thr <- map_dfr(threshold_grid, function(th) {
    pred <- if_else(preds_tbl$prob >= th, "yes", "no") |> factor(levels = c("no", "yes"))
    tibble(threshold = th, f2 = yardstick::f_meas_vec(preds_tbl$truth, pred, beta = 2))
  })

  best_row <- f_by_thr[which.max(f_by_thr$f2), , drop = FALSE]
  best_row$threshold
}

tuned_thresholds <- imap(finalized_clf, function(wf, name) {
  cat("[Thresholds] Optimizing decision threshold for:", name, "\n")
  thr <- compute_best_threshold(wf)
  cat("[Thresholds] Best threshold for", name, "=", round(thr, 3), "\n")
  thr
})

# Inspect tuned thresholds (optional)
print(tuned_thresholds)
```

## Final fit and test evaluation (best pipeline by combined RMSE)

```{r}
#| label: final-test

best_pipeline <- val_summary$pipeline[1]
best_clf_name <- val_summary$clf[1]
best_reg_name <- val_summary$reg[1]

cat("Best pipeline:", best_pipeline, "=", best_clf_name, "+", best_reg_name, "\n")

# Finalize on full pretest
final_clf_fit <- fit(finalized_clf[[best_clf_name]], data = pretest_df)

# Severity final fit on positives only (pretest)
pretest_pos <- pretest_df |>
  filter(strike_count > 0)
final_reg_fit <- fit(finalized_reg[[best_reg_name]], data = pretest_pos)

# Test predictions with tuned threshold for the selected classifier
best_threshold <- tuned_thresholds[[best_clf_name]]
test_prob_yes <- predict(final_clf_fit, new_data = test_df, type = "prob")$.pred_yes
test_prob_yes[is.na(test_prob_yes)] <- 0
test_prob_yes <- pmin(pmax(test_prob_yes, 0), 1)
test_occ_pred <- if_else(test_prob_yes >= best_threshold, "yes", "no") |> factor(levels = c("no", "yes"))
truth_occ <- factor(test_df$y_occ, levels = c("no","yes"))
pr_auc_val <- yardstick::pr_auc_vec(truth_occ, test_prob_yes, event_level = "second")
roc_auc_val <- yardstick::roc_auc_vec(truth_occ, test_prob_yes, event_level = "second")
f2_val <- yardstick::f_meas_vec(truth_occ, test_occ_pred, beta = 2)
test_occ <- tibble::tibble(
  .metric = c("pr_auc", "roc_auc", "f_meas2"),
  .estimator = "binary",
  .estimate = c(pr_auc_val, roc_auc_val, f2_val)
)

test_sev_hat <- predict_severity(final_reg_fit, new_data = test_df, model_name = best_reg_name)
test_sev_hat <- pmax(pmin(test_sev_hat, quantile(test_sev_hat, 0.999, na.rm = TRUE)), 0)
test_sev_hat[!is.finite(test_sev_hat)] <- 0
test_sev <- metric_set(mae, rmse, rsq)(
  bind_cols(truth = test_df$strike_count[test_df$strike_count > 0],
            .pred = test_sev_hat[test_df$strike_count > 0]),
  truth = truth, estimate = .pred
)

test_combined_hat <- test_prob_yes * test_sev_hat
test_combined <- metric_set(mae, rmse, rsq)(
  bind_cols(truth = test_df$strike_count, .pred = test_combined_hat),
  truth = truth, estimate = .pred
)

cat("\nTest occurrence metrics:\n"); print(test_occ)
cat("\nTest severity metrics (on positives):\n"); print(test_sev)
cat("\nTest combined metrics (all weeks):\n"); print(test_combined)

# Naive baselines for reference
naive_zero <- tibble(.metric = c("mae","rmse","rsq"), .estimator = "standard",
                     .estimate = c(mae_vec(test_df$strike_count, rep(0, nrow(test_df))),
                                   rmse_vec(test_df$strike_count, rep(0, nrow(test_df))),
                                   rsq_vec(test_df$strike_count, rep(0, nrow(test_df)))))

combined_counts <- c(pretest_df$strike_count, test_df$strike_count)
lagged_all <- dplyr::lag(combined_counts, 1)
lag1_preds <- tail(lagged_all, nrow(test_df))
lag1_preds[is.na(lag1_preds)] <- 0
naive_lag1 <- tibble(.metric = c("mae","rmse","rsq"), .estimator = "standard",
                     .estimate = c(mae_vec(test_df$strike_count, lag1_preds),
                                   rmse_vec(test_df$strike_count, lag1_preds),
                                   rsq_vec(test_df$strike_count, lag1_preds)))

roll4 <- zoo::rollapplyr(c(pretest_df$strike_count, test_df$strike_count), 4, mean, partial = TRUE)
roll4_test <- tail(roll4, nrow(test_df))
naive_roll4 <- tibble(.metric = c("mae","rmse","rsq"), .estimator = "standard",
                      .estimate = c(mae_vec(test_df$strike_count, roll4_test),
                                    rmse_vec(test_df$strike_count, roll4_test),
                                    rsq_vec(test_df$strike_count, roll4_test)))

cat("\nNaive baselines (all weeks):\n")
print(list(zero = naive_zero, lag1 = naive_lag1, roll4 = naive_roll4))
```

## Save outputs

```{r}
#| label: save-outputs

dir.create("model_outputs", showWarnings = FALSE)

# 1) Validation summaries and per-fold metrics
readr::write_csv(val_summary, "../../outputs/models/two_stage_val_summary.csv")

val_folds_tbl <- bind_rows(pipeline_results, .id = "pipeline")
readr::write_csv(val_folds_tbl, "../../outputs/models/two_stage_val_folds.csv")

# 2) Final model objects
saveRDS(final_clf_fit, "../../outputs/models/two_stage_final_classifier.rds")
saveRDS(final_reg_fit, "../../outputs/models/two_stage_final_regressor.rds")

# 3) Test predictions and metrics
test_predictions_tbl <- test_df |>
  select(date, action_geo_adm1_code, action_geo_country_code, strike_count, y_occ) |>
  mutate(prob_yes = test_prob_yes,
         sev_hat = test_sev_hat,
         expected_count = prob_yes * sev_hat)

arrow::write_parquet(test_predictions_tbl, "../../outputs/models/two_stage_test_predictions.parquet")

test_metrics_tbl <- bind_rows(
  test_occ |> mutate(stage = "occurrence"),
  test_sev |> mutate(stage = "severity_positive"),
  test_combined |> mutate(stage = "combined_all")
)
readr::write_csv(test_metrics_tbl, "../../outputs/models/two_stage_test_metrics.csv")

# 4) Best pipeline record
best_info <- tibble(
  best_pipeline = best_pipeline,
  classifier = best_clf_name,
  regressor = best_reg_name
)
readr::write_csv(best_info, "../../outputs/models/two_stage_best_pipeline.csv")

# 5) Save tuned threshold
thr_tbl <- tibble(classifier = best_clf_name, tuned_threshold = best_threshold)
readr::write_csv(thr_tbl, "../../outputs/models/two_stage_tuned_threshold.csv")

# 6) Save naive baselines
naive_tbl <- tibble(
  baseline = c("zero","lag1","roll4"),
  mae = c(naive_zero$.estimate[1], naive_lag1$.estimate[1], naive_roll4$.estimate[1]),
  rmse = c(naive_zero$.estimate[2], naive_lag1$.estimate[2], naive_roll4$.estimate[2]),
  rsq = c(naive_zero$.estimate[3], naive_lag1$.estimate[3], naive_roll4$.estimate[3])
)
readr::write_csv(naive_tbl, "../../outputs/models/two_stage_naive_baselines.csv")

# 7) Save compact RDS summary of validation and final test results
results_rds <- list(
  val_summary = val_summary,
  best_info = best_info,
  tuned_thresholds = tuned_thresholds,
  test_metrics = list(
    occurrence = test_occ,
    severity_positive = test_sev,
    combined_all = test_combined
  )
)
saveRDS(results_rds, "../../outputs/models/two_stage_results_summary.rds")
```

## Feature importance for final models

```{r}
#| label: feature-importance

# Try model-based importance first; fall back silently if unsupported
clf_engine <- try(workflows::extract_fit_parsnip(final_clf_fit)$fit, silent = TRUE)
reg_engine <- try(workflows::extract_fit_parsnip(final_reg_fit)$fit, silent = TRUE)

clf_vip_plot <- try(vip::vip(clf_engine, num_features = 20), silent = TRUE)
reg_vip_plot <- try(vip::vip(reg_engine, num_features = 20), silent = TRUE)

if (!inherits(clf_vip_plot, "try-error")) {
  ggsave("../../outputs/models/two_stage_feature_importance_classifier.png", clf_vip_plot, width = 9, height = 6, dpi = 300)
}
if (!inherits(reg_vip_plot, "try-error")) {
  ggsave("../../outputs/models/two_stage_feature_importance_regressor.png", reg_vip_plot, width = 9, height = 6, dpi = 300)
}
```

## Appendix: single-stage severity models (reference)

```{r}
#| label: appendix-single-stage

# Train on pretest (all weeks), evaluate on test

tree_recipe <- recipe(strike_count ~ ., data = pretest_df) |>
  update_role(date, new_role = "id") |>
  step_novel(action_geo_adm1_code, action_geo_country_code) |>
  step_other(action_geo_adm1_code, threshold = 0.01) |>
  step_dummy(action_geo_adm1_code, action_geo_country_code) |>
  step_mutate(is_year_end = as.numeric(is_year_end)) |>
  step_dummy(all_nominal_predictors()) |>
  step_impute_median(all_numeric_predictors()) |>
  step_zv(all_predictors())

single_specs <- list(
  lm_log = linear_reg() |> set_engine("lm") |> set_mode("regression"),
  xgb_poi = xgb_poi_spec,
  lgb_poi = lgb_poi_spec,
  lgb_twd = lgb_twd_spec
)

single_wfs <- list(
  lm_log = workflow() |> add_recipe(tree_recipe |> step_mutate(log_count = log1p(strike_count)) |> step_rm(strike_count)) |> add_model(single_specs$lm_log),
  xgb_poi = workflow() |> add_recipe(tree_recipe) |> add_model(single_specs$xgb_poi),
  lgb_poi = workflow() |> add_recipe(tree_recipe) |> add_model(single_specs$lgb_poi),
  lgb_twd = workflow() |> add_recipe(tree_recipe) |> add_model(single_specs$lgb_twd)
)

single_fit_predict <- function(wf, name) {
  fit_obj <- fit(wf, data = pretest_df)
  preds <- predict(fit_obj, new_data = test_df)$.pred
  # Back-transform for lm_log: recipe models log1p(strike_count) as outcome
  # so we invert with expm1() to return to the original count scale.
  if (name == "lm_log") preds <- pmax(expm1(preds), 0)
  tibble(model = name, mae = mae_vec(test_df$strike_count, preds), rmse = rmse_vec(test_df$strike_count, preds), r2 = rsq_vec(test_df$strike_count, preds))
}

single_results <- imap_dfr(single_wfs, single_fit_predict)
print(single_results)

# Save appendix results
readr::write_csv(single_results, "../../outputs/models/single_stage_results.csv")
```


