---
title: "Single-Stage Baselines (Tweedie/Poisson/LM log1p)"
format: html
---

**Metrics Implementation Notes:**

- **Primary selection metrics**: MAE, RMSE, R² on original scale (validation uses pooled/micro-averaged metrics over all validation rows)
- **Zero/non-zero reporting**: Computed from concatenated validation predictions (pooled) for each model
- **Log-scale models**: Report log-scale R² as supplementary (uses saved `.pred_raw`); selection uses original-scale metrics
- **Smearing correction**: Duan's smearing is applied consistently for log models in validation (per-fold factors from analysis residuals) and test (single factor from pretest residuals)
- **Poisson/Tweedie GLMs**: Pseudo-R² (deviance-based) reported as supplementary
- **Model selection**: Toggle between RMSE or MAE via `selection_metric`

## Setup

```{r}
#| label: setup
#| message: false

library(tidymodels)
library(tidyverse)
library(lubridate)
library(arrow)
library(bonsai)
library(finetune)
library(glmnet)
library(janitor)
library(doFuture)
library(vip)

# Set tidymodels and conflict preferences
tidymodels_prefer()
conflicted::conflicts_prefer(dplyr::lag)
conflicted::conflicts_prefer(stringr::fixed)

# Set parallel-safe RNG FIRST
RNGkind("L'Ecuyer-CMRG")

# Set base seed
set.seed(2024)

# Parallel backend setup
registerDoFuture()
plan(multisession, workers = 4)
options(future.globals.maxSize = 8 * 1024^3)

# Load helper functions
source("../../helpers/helpers.R")

# Prediction capping controls
cap_predictions <- FALSE   # set TRUE to enable winsorization
cap_quantile <- 0.999      # cap at this quantile of training response

# Model selection metric toggle
selection_metric <- "rmse"  # "rmse" or "mae" - metric used for model selection

# Validation feature importance controls
save_val_importance <- FALSE  # Skip slow permutation importance
perm_importance_nsim <- 5
```

## Data and splits 

```{r}
#| label: data

adm_full <- read_parquet("../../data/analysis/adm_week_full.parquet") |>
  clean_names()

predictor_cols <- adm_full |>
  select(
    action_geo_adm1_code, action_geo_country_code,
    time_trend, year_sin, year_cos, month, quarter, is_year_end,
    strike_count_lag1, strike_count_lag2, strike_count_lag4,
    rolling_avg_4wk, rolling_avg_8wk,
    total_articles_lag1, total_articles_lag2,
    total_mentions_lag1, total_mentions_lag2,
    avg_tone_lag1, avg_tone_lag2,
    actor_diversity_lag1,
    unique_actor1_types_lag1, unique_actor2_types_lag1,
    prop_gov_lag1, prop_labor_lag1, prop_civil_lag1,
    country_strikes_lag1, contig_strikes_t1, distw_strikes_t1
  ) |>
  names()

model_df <- adm_full |>
  select(date, strike_count, all_of(predictor_cols)) |>
  arrange(date)

all_weeks <- sort(unique(model_df$date))
test_weeks <- tail(all_weeks, 52)
pretest_df <- model_df |> filter(date < min(test_weeks))
test_df <- model_df |> filter(date %in% test_weeks)

assessment_weeks <- 26
n_folds <- 3
pretest_max <- max(pretest_df$date)
assess_end_dates <- pretest_max - weeks(rev(seq(0, by = assessment_weeks, length.out = n_folds))) ## ?
assess_start_dates <- assess_end_dates - weeks(assessment_weeks - 1)

make_time_split <- function(df, start_date, end_date) {
  analysis_idx <- which(df$date < start_date)
  assessment_idx <- which(df$date >= start_date & df$date <= end_date)
  rsample::make_splits(list(analysis = analysis_idx, assessment = assessment_idx), df) ## ?
}

val_splits <- map2(assess_start_dates, assess_end_dates, ~ make_time_split(pretest_df, .x, .y)) ## ?
val_resamples <- rsample::manual_rset(val_splits, paste0("fold", seq_along(val_splits))) ## ?
```

## Main recipe

```{r}
#| label: recipe

tree_recipe <- recipe(strike_count ~ ., data = pretest_df) |>
  update_role(date, new_role = "id") |>
  step_novel(action_geo_adm1_code, action_geo_country_code) |>
  step_other(action_geo_adm1_code, threshold = 0.02) |>
  step_mutate(is_year_end = as.numeric(is_year_end)) |>
  step_dummy(all_nominal_predictors()) |>
  step_impute_median(all_numeric_predictors()) |>
  step_lincomb(all_numeric_predictors()) |>
  step_corr(all_numeric_predictors(), threshold = 0.999) |>
  step_zv(all_predictors())
```

## Model specifications

```{r}
#| label: models

# Make racing metric match selection_metric to avoid mismatch warnings
if (identical(selection_metric, "rmse")) {
  reg_metrics <- metric_set(rmse, mae, rsq)
} else if (identical(selection_metric, "mae")) {
  reg_metrics <- metric_set(mae, rmse, rsq)
} else {
  stop("selection_metric must be 'rmse' or 'mae'")
}

# LM on log1p(count)
lm_log_spec <- linear_reg() |> set_engine("lm") |> set_mode("regression")
lm_log_recipe <- tree_recipe |>
  step_log(strike_count, offset = 1, skip = TRUE)

# GLMNET (log1p target)
glmnet_log_spec <- linear_reg(
  penalty = tune(),
  mixture = tune()
) |>
  set_engine("glmnet") |>
  set_mode("regression")

# Base LightGBM specification
base_lgbm <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_mode("regression")

# LightGBM Tweedie (p=1.3 based on earlier validation results)
lgbm_twd_spec <- base_lgbm |> set_engine("lightgbm", objective = "tweedie", tweedie_variance_power = 1.3, num_threads = 1)

# LightGBM Poisson
lgbm_poi_spec <- base_lgbm |> set_engine("lightgbm", objective = "poisson", num_threads = 1)

# LightGBM L2 (log1p target)
lgbm_l2_spec <- base_lgbm |> set_engine("lightgbm", objective = "regression", num_threads = 1)

# Base XGBoost specification
base_xgb <- boost_tree(
  trees = tune(), tree_depth = tune(), min_n = tune(),
  loss_reduction = tune(), sample_size = tune(), mtry = tune(), learn_rate = tune()
) |>
  set_mode("regression")

# XGBoost Poisson
xgb_poi_spec <- base_xgb |> set_engine("xgboost", objective = "count:poisson", tree_method = "hist", nthread = 1)

# XGBoost L2 (log1p target)
xgb_l2_spec <- base_xgb |> set_engine("xgboost", objective = "reg:squarederror", tree_method = "hist", nthread = 1)

# Random Forest (log1p target)
rf_log_spec <- rand_forest(
  trees = tune(), mtry = tune(), min_n = tune()
) |>
  set_engine("ranger", num.threads = 1) |>
  set_mode("regression")
```

## Tuning grids

```{r}
#| label: grids

# LightGBM/XGBoost tuning grid
grid_boosted <- grid_space_filling(    ## study these parameters
  trees(range = c(200, 600)),
  tree_depth(range = c(3, 8)),
  min_n(range = c(2, 20)),
  loss_reduction(range = c(0, 5)),
  sample_size = sample_prop(range = c(0.6, 0.9)),
  mtry(range = c(2, min(12, length(predictor_cols)))),
  learn_rate(range = c(-3, -1), trans = log10_trans()),
  size = 24
)

# Random Forest tuning grid
grid_rf <- grid_space_filling(
  trees(range = c(300, 1000)),
  min_n(range = c(2, 20)),
  mtry(range = c(2, min(12, length(predictor_cols)))),
  size = 20
)

# GLMNET tuning grid
grid_glmnet <- grid_regular(
  penalty(range = c(-5, 0), trans = log10_trans()),
  mixture(range = c(0, 1)),
  levels = c(penalty = 10, mixture = 5)  # Named vector: 10 × 5 = 50 configs
)
```

## Workflows

```{r}
#| label: workflows

# LM log1p 
wf_lm_log <- workflow() |> add_model(lm_log_spec) |> add_recipe(lm_log_recipe)

# GLMNET log1p
wf_glmnet_log <- workflow() |> add_model(glmnet_log_spec) |> add_recipe(lm_log_recipe)

# LightGBM Poisson
wf_lgbm_poi <- workflow() |> add_model(lgbm_poi_spec) |> add_recipe(tree_recipe)

# LightGBM Tweedie
wf_lgbm_twd <- workflow() |> add_model(lgbm_twd_spec) |> add_recipe(tree_recipe)  

# LightGBM L2 (log1p target)
wf_lgbm_log <- workflow() |> add_model(lgbm_l2_spec) |> add_recipe(lm_log_recipe)

# XGBoost Poisson
wf_xgb_poi <- workflow() |> add_model(xgb_poi_spec) |> add_recipe(tree_recipe)

# XGBoost L2 (log1p target)
wf_xgb_log <- workflow() |> add_model(xgb_l2_spec) |> add_recipe(lm_log_recipe)
```

## Tuning and validation

```{r}
#| label: tuning
#| message: true
#| warning: true

# record start time
start_time <- Sys.time()

# Model selection toggles - edit this list to run only specific models
models_to_run <- c(
  "lm_log",      # Linear model on log1p
  "glmnet_log",  # GLMNET on log1p
  "lgbm_poi",    # LightGBM Poisson
  "lgbm_twd",    # LightGBM Tweedie
  "lgbm_log",    # LightGBM L2 on log1p
  "xgb_poi",     # XGBoost Poisson
  "xgb_log"      # XGBoost L2 on log1p
)

# Standard grid control (no racing)
ctrl_grid <- control_grid(
  save_pred = FALSE, verbose = TRUE, allow_par = TRUE,
  parallel_over = "resamples", save_workflow = FALSE
)  

# Print warnings as they occur during tuning
old_warn <- getOption("warn"); options(warn = 1)

# Containers for fitted models, validation results, and predictions
fitted_models <- list()
val_results <- list()
val_predictions <- list()

# LM log 1p
if ("lm_log" %in% models_to_run) {
  cat("Prepping LM log model for validation...\n")
  final_lm_log <- wf_lm_log  # No tuning required, just feed workflow to model object
  # Refit and collect validation metrics
  cat("Fitting LM log model across validation folds...\n")
  lm_log_val <- collect_fold_metrics(
    final_lm_log,
    val_resamples$splits,
    backtransform = function(x) pmax(expm1(x), 0),
    model_type = "log",
    use_smearing = TRUE,
    split_ids = val_resamples$id,
    model_name = "lm_log",
    capture_last_fit = TRUE
  )
  # Get the last fold (full pretest data)
  fitted_models$lm_log <- lm_log_val$last_fit
  # Collect validation results
  val_results$lm_log <- lm_log_val$metrics
  val_predictions$lm_log <- lm_log_val$predictions
} else {
  cat(" Skipping LM log model\n")       
  final_lm_log <- NULL
}

# GLMNET on log1p
if ("glmnet_log" %in% models_to_run) {
  cat(sprintf(" Tuning GLMNET (log1p)... (grid=%d, folds=%d)\n", nrow(grid_glmnet), nrow(val_resamples)))
  cat(" Grid check - penalty levels:", length(unique(grid_glmnet$penalty)), "\n")
  cat(" Grid check - mixture levels:", length(unique(grid_glmnet$mixture)), "\n")
  cat(" Grid check - total configs:", nrow(grid_glmnet), "\n")
  # Tuning with standard grid (no racing)
  tuned_glmnet_log <- tune_grid(wf_glmnet_log, resamples = val_resamples, grid = grid_glmnet, metrics = reg_metrics, control = ctrl_grid)
  # Persist tuning metrics and report evaluated counts
  met_glmnet_log <- collect_metrics(tuned_glmnet_log)
  readr::write_csv(met_glmnet_log, sprintf("../../outputs/models/sst_tuning_metrics_glmnet_log__%s.csv", selection_metric))
  cat(" Post-tuning check - unique configs evaluated:", length(unique(met_glmnet_log$.config)), "of", nrow(grid_glmnet), "\n")
  # Finalize workflow
  final_glmnet_log <- finalize_workflow(wf_glmnet_log, select_best(tuned_glmnet_log, metric = selection_metric))
  # Refit and collect validation metrics
  cat("Re-fitting finalized model across validation folds for derived metrics...\n")
  glmnet_log_val <- collect_fold_metrics(
    final_glmnet_log,
    val_resamples$splits,
    backtransform = function(x) pmax(expm1(x), 0),
    model_type = "log",
    use_smearing = TRUE,
    split_ids = val_resamples$id,
    model_name = "glmnet_log",
    capture_last_fit = TRUE
  )
  # Extract the fitted model from the last fold
  fitted_models$glmnet_log <- glmnet_log_val$last_fit
  # Collect validation results
  val_results$glmnet_log <- glmnet_log_val$metrics
  val_predictions$glmnet_log <- glmnet_log_val$predictions
  cat("Done GLMNET (log1p).\n")
} else {
  cat(" Skipping GLMNET (log1p)\n")
  tuned_glmnet_log <- NULL
  final_glmnet_log <- NULL
}

# LightGBM Poisson
if ("lgbm_poi" %in% models_to_run) {
  cat(sprintf("Tuning LightGBM Poisson... (grid=%d, folds=%d)\n", nrow(grid_boosted), nrow(val_resamples)))
  # Tuning with standard grid (no racing)
  tuned_lgbm_poi <- tune_grid(wf_lgbm_poi, resamples = val_resamples, grid = grid_boosted, metrics = reg_metrics, control = ctrl_grid)
  # Persist tuning metrics and report evaluated counts
  met_lgbm_poi <- collect_metrics(tuned_lgbm_poi)
  readr::write_csv(met_lgbm_poi, sprintf("../../outputs/models/sst_tuning_metrics_lgbm_poi__%s.csv", selection_metric))
  cat(" Post-tuning check - unique configs evaluated:", length(unique(met_lgbm_poi$.config)), "of", nrow(grid_boosted), "\n")
  # Finalize workflow
  final_lgbm_poi <- finalize_workflow(wf_lgbm_poi, select_best(tuned_lgbm_poi, metric = selection_metric))
  # Refit and collect validation metrics
  cat("Re-fitting finalized model across validation folds for derived metrics...\n")
  lgbm_poi_val <- collect_fold_metrics(
    final_lgbm_poi,
    val_resamples$splits,
    backtransform = identity,
    model_type = "poisson",
    split_ids = val_resamples$id,
    model_name = "lgbm_poi",
    capture_last_fit = TRUE
  )
  # Extract the fitted model from the last fold
  fitted_models$lgbm_poi <- lgbm_poi_val$last_fit
  # Collect validation results
  val_results$lgbm_poi <- lgbm_poi_val$metrics
  val_predictions$lgbm_poi <- lgbm_poi_val$predictions
  cat("Done LightGBM Poisson.\n")
} else {
  cat(" Skipping LightGBM Poisson\n")
  tuned_lgbm_poi <- NULL
  final_lgbm_poi <- NULL
}

# LightGBM Tweedie
if ("lgbm_twd" %in% models_to_run) {
  cat(sprintf("Tuning LightGBM Tweedie (p=1.3)... (grid=%d, folds=%d)\n", nrow(grid_boosted), nrow(val_resamples)))
  # Tuning with standard grid (no racing)
  tuned_lgbm_twd <- tune_grid(wf_lgbm_twd, resamples = val_resamples, grid = grid_boosted, metrics = reg_metrics, control = ctrl_grid)
  # Persist tuning metrics and report evaluated counts
  met_lgbm_twd <- collect_metrics(tuned_lgbm_twd)
  readr::write_csv(met_lgbm_twd, sprintf("../../outputs/models/sst_tuning_metrics_lgbm_twd__%s.csv", selection_metric))
  cat(" Post-tuning check - unique configs evaluated:", length(unique(met_lgbm_twd$.config)), "of", nrow(grid_boosted), "\n")
  # Finalize workflow
    final_lgbm_twd <- finalize_workflow(wf_lgbm_twd, select_best(tuned_lgbm_twd, metric = selection_metric))
  # Refit and collect validation metrics
  cat("Re-fitting finalized model across validation folds for derived metrics...\n")
  lgbm_twd_val <- collect_fold_metrics(
    final_lgbm_twd,
    val_resamples$splits,
    backtransform = identity,
    model_type = "tweedie",
    split_ids = val_resamples$id,
    model_name = "lgbm_twd",
    capture_last_fit = TRUE
  )
  # Extract the fitted model from the last fold
  fitted_models$lgbm_twd <- lgbm_twd_val$last_fit
  # Collect validation results
  val_results$lgbm_twd <- lgbm_twd_val$metrics
  val_predictions$lgbm_twd <- lgbm_twd_val$predictions
  cat("Done LightGBM Tweedie.\n")
} else {
  cat(" Skipping LightGBM Tweedie\n")
  tuned_lgbm_twd <- NULL
  final_lgbm_twd <- NULL
}

# LightGBM L2 (log1p target)
if ("lgbm_log" %in% models_to_run) {
  cat(sprintf("Tuning LightGBM (L2) on log1p... (grid=%d, folds=%d)\n", nrow(grid_boosted), nrow(val_resamples)))
  # Tuning with standard grid (no racing)
  tuned_lgbm_log <- tune_grid(wf_lgbm_log, resamples = val_resamples, grid = grid_boosted, metrics = reg_metrics, control = ctrl_grid)
  # Persist tuning metrics and report evaluated counts
  met_lgbm_log <- collect_metrics(tuned_lgbm_log)
  readr::write_csv(met_lgbm_log, sprintf("../../outputs/models/sst_tuning_metrics_lgbm_log__%s.csv", selection_metric))
  cat(" Post-tuning check - unique configs evaluated:", length(unique(met_lgbm_log$.config)), "of", nrow(grid_boosted), "\n")
  # Finalize workflow
  final_lgbm_log <- finalize_workflow(wf_lgbm_log, select_best(tuned_lgbm_log, metric = selection_metric))
  # Refit and collect validation metrics
  cat("Re-fitting finalized model across validation folds for derived metrics...\n")
  lgbm_log_val <- collect_fold_metrics(
    final_lgbm_log,
    val_resamples$splits,
    backtransform = function(x) pmax(expm1(x), 0),
    model_type = "log",
    use_smearing = TRUE,
    split_ids = val_resamples$id,
    model_name = "lgbm_log",
    capture_last_fit = TRUE
  )
  # Extract the fitted model from the last fold
  fitted_models$lgbm_log <- lgbm_log_val$last_fit
  # Collect validation results
  val_results$lgbm_log <- lgbm_log_val$metrics
  val_predictions$lgbm_log <- lgbm_log_val$predictions
  cat(" Done LightGBM (L2) on log1p.\n")
} else {
  cat(" Skipping LightGBM (L2) on log1p\n")
  tuned_lgbm_log <- NULL
  final_lgbm_log <- NULL
}

# XGBoost Poisson
if ("xgb_poi" %in% models_to_run) {
  cat(sprintf("Tuning XGBoost Poisson... (grid=%d, folds=%d)\n", nrow(grid_boosted), nrow(val_resamples)))
  # Tuning with standard grid (no racing)
  tuned_xgb_poi <- tune_grid(wf_xgb_poi, resamples = val_resamples, grid = grid_boosted, metrics = reg_metrics, control = ctrl_grid)
  # Persist tuning metrics and report evaluated counts
  met_xgb_poi <- collect_metrics(tuned_xgb_poi)
  readr::write_csv(met_xgb_poi, sprintf("../../outputs/models/sst_tuning_metrics_xgb_poi__%s.csv", selection_metric))
  cat(" Post-tuning check - unique configs evaluated:", length(unique(met_xgb_poi$.config)), "of", nrow(grid_boosted), "\n")
  # Finalize workflow
  final_xgb_poi <- finalize_workflow(wf_xgb_poi, select_best(tuned_xgb_poi, metric = selection_metric))
  # Refit and collect validation metrics
  cat("Re-fitting finalized model across validation folds for derived metrics...\n")
  xgb_poi_val <- collect_fold_metrics(
    final_xgb_poi,
    val_resamples$splits,
    backtransform = identity,
    model_type = "poisson",
    split_ids = val_resamples$id,
    model_name = "xgb_poi",
    capture_last_fit = TRUE
  )
  # Extract the fitted model from the last fold
  fitted_models$xgb_poi <- xgb_poi_val$last_fit
  # Collect validation results
  val_results$xgb_poi <- xgb_poi_val$metrics
  val_predictions$xgb_poi <- xgb_poi_val$predictions
  cat("Done XGBoost Poisson.\n")
} else {
  cat("Skipping XGBoost Poisson\n")
  tuned_xgb_poi <- NULL
  final_xgb_poi <- NULL
}

# XGBoost L2 (log1p target)
if ("xgb_log" %in% models_to_run) {
  cat(sprintf("Tuning XGBoost (L2) on log1p... (grid=%d, folds=%d)\n", nrow(grid_boosted), nrow(val_resamples)))
  # Tuning with standard grid (no racing)
  tuned_xgb_log <- tune_grid(wf_xgb_log, resamples = val_resamples, grid = grid_boosted, metrics = reg_metrics, control = ctrl_grid)
  # Persist tuning metrics and report evaluated counts
  met_xgb_log <- collect_metrics(tuned_xgb_log)
  readr::write_csv(met_xgb_log, sprintf("../../outputs/models/sst_tuning_metrics_xgb_log__%s.csv", selection_metric))
  cat(" Post-tuning check - unique configs evaluated:", length(unique(met_xgb_log$.config)), "of", nrow(grid_boosted), "\n")
  # Finalize workflow
  final_xgb_log <- finalize_workflow(wf_xgb_log, select_best(tuned_xgb_log, metric = selection_metric))
  # Refit and collect validation metrics
  cat("Re-fitting finalized model across validation folds for derived metrics...\n")
  xgb_log_val <- collect_fold_metrics(
    final_xgb_log,
    val_resamples$splits,
    backtransform = function(x) pmax(expm1(x), 0),
    model_type = "log",
    use_smearing = TRUE,
    split_ids = val_resamples$id,
    model_name = "xgb_log",
    capture_last_fit = TRUE
  )
  # Extract the fitted model from the last fold
  fitted_models$xgb_log <- xgb_log_val$last_fit
  # Collect validation results
  val_results$xgb_log <- xgb_log_val$metrics
  val_predictions$xgb_log <- xgb_log_val$predictions
  cat("Done XGBoost (L2) on log1p.\n")
} else {
  cat("Skipping XGBoost (L2) on log1p\n")
  tuned_xgb_log <- NULL
  final_xgb_log <- NULL
}

# Summarize validation results
if (length(val_results) > 0) {
  val_results <- val_results |>
    dplyr::bind_rows(.id = "model")

  # Already pooled per compute_metrics_from_preds; just arrange
  val_summary <- val_results |>
    dplyr::arrange(.data[[selection_metric]])

  cat(sprintf("Validation summary (ranked by %s):\n", toupper(selection_metric))); print(val_summary)
} else {
  cat("No models were run. Check models_to_run list.\n")
  val_summary <- data.frame(model = character(0), mae = numeric(0), rmse = numeric(0), rsq = numeric(0))
}

# Save validation results for later analysis (metric-aware filenames)
readr::write_csv(val_results, sprintf("../../outputs/models/sst_val_results_detailed_%s.csv", selection_metric))
readr::write_csv(val_summary, sprintf("../../outputs/models/sst_val_summary_%s.csv", selection_metric))

# Save all fitted models for later analysis
saveRDS(fitted_models, sprintf("../../outputs/models/sst_all_fitted_models_%s.rds", selection_metric))

# record end time and print elapsed time
end_time <- Sys.time()
elapsed <- end_time - start_time
cat("Time elapsed:", elapsed, "\n")

# Restore warning option
options(warn = old_warn)         
```

## Validation: Zero vs Non-Zero Metrics

```{r}
#| label: validation-separate-metrics

if (length(val_predictions) > 0) {
  cat("\n[Validation] Computing zero/non-zero and supplementary metrics from saved predictions...\n")
  
  val_separate_metrics <- map_dfr(names(val_predictions), function(model_name) {
    preds_df <- val_predictions[[model_name]]
    model_type <- unique(preds_df$model_type)[1]
    
    # Separate by zero/non-zero (pooled on concatenated predictions)
    zero_mask <- preds_df$truth == 0
    nonzero_mask <- preds_df$truth > 0
    
    basic_metrics <- bind_rows(
      tibble(
        model = model_name,
        week_type = "zero",
        .metric = c("mae", "rmse"),
        .estimate = c(
          yardstick::mae_vec(preds_df$truth[zero_mask], preds_df$.pred[zero_mask]),
          yardstick::rmse_vec(preds_df$truth[zero_mask], preds_df$.pred[zero_mask])
        )
      ),
      tibble(
        model = model_name,
        week_type = "non_zero",
        .metric = c("mae", "rmse", "rsq"),
        .estimate = c(
          yardstick::mae_vec(preds_df$truth[nonzero_mask], preds_df$.pred[nonzero_mask]),
          yardstick::rmse_vec(preds_df$truth[nonzero_mask], preds_df$.pred[nonzero_mask]),
          yardstick::rsq_vec(preds_df$truth[nonzero_mask], preds_df$.pred[nonzero_mask])
        )
      )
    )
    
    # Add supplementary metrics (overall, not by zero/non-zero)
    supp_metrics <- tibble()
    if (model_type %in% c("poisson", "tweedie")) {
      pseudo_r2 <- compute_pseudo_r2(preds_df$truth, preds_df$.pred, family = model_type)
      supp_metrics <- bind_rows(supp_metrics,
        tibble(model = model_name, week_type = "overall", .metric = "pseudo_r2", .estimate = pseudo_r2)
      )
    }
    if (model_type == "log" && ".pred_raw" %in% names(preds_df)) {
      r2_log <- yardstick::rsq_vec(log1p(preds_df$truth), preds_df$.pred_raw)
      supp_metrics <- bind_rows(supp_metrics,
        tibble(model = model_name, week_type = "overall", .metric = "r2_log_scale", .estimate = r2_log)
      )
    }
    
    bind_rows(basic_metrics, supp_metrics)
  })
  
  cat("\n[Validation] Zero/Non-zero and supplementary metrics by model:\n")
  print(val_separate_metrics, n = 24)
  
  readr::write_csv(val_separate_metrics, sprintf("../../outputs/models/sst_val_separate_metrics_%s.csv", selection_metric))
} else {
  cat("\n[Validation] No predictions saved.")
}
```

## Final fit on pretest and test evaluation

```{r}
#| label: final-test

best_model <- val_summary$model[1]
cat(" Best single-stage model:", best_model, "\n")

final_wf <- switch(best_model,
  lm_log = final_lm_log,
  lgbm_poi = final_lgbm_poi,
  xgb_poi = final_xgb_poi,
  lgbm_log = final_lgbm_log,
  xgb_log = final_xgb_log,
  lgbm_twd = final_lgbm_twd,
  glmnet_log = final_glmnet_log
)

final_fit <- fit(final_wf, data = pretest_df)
test_pred <- predict(final_fit, new_data = test_df)$.pred

# Back-transform for log1p models
if (best_model %in% c("lm_log", "lgbm_log", "xgb_log", "glmnet_log")) {
  # Apply smearing correction on test
  test_pred_log <- predict(final_fit, new_data = test_df)$.pred
  train_pred_log <- predict(final_fit, new_data = pretest_df)$.pred
  log_resids <- log1p(pretest_df$strike_count) - train_pred_log
  smear_factor <- compute_smearing_factor(log_resids)
  test_pred <- pmax(smearing_backtransform(test_pred_log, smear_factor) - 1, 0)
}

# Optional: cap using training (pretest) response distribution to avoid leakage
if (isTRUE(cap_predictions)) {
  cap_val <- stats::quantile(pretest_df$strike_count, cap_quantile, na.rm = TRUE)
  test_pred <- pmin(test_pred, cap_val)
}

# Enforce non-negativity and clean up
test_pred <- pmax(test_pred, 0)
test_pred[!is.finite(test_pred)] <- 0

test_tbl <- metric_set(mae, rmse, rsq)(
  bind_cols(truth = test_df$strike_count, .pred = test_pred),
  truth = truth, estimate = .pred
)

# Add supplementary metrics based on model type
supplementary_metrics <- tibble()

if (best_model %in% c("lgbm_poi", "xgb_poi")) {
  # Pseudo-R2 for Poisson models
  pseudo_r2_val <- compute_pseudo_r2(test_df$strike_count, test_pred, family = "poisson")
  supplementary_metrics <- bind_rows(supplementary_metrics,
    tibble(.metric = "pseudo_r2_poisson", .estimator = "standard", .estimate = pseudo_r2_val))
  cat(sprintf(" Pseudo-R² (Poisson deviance): %.4f\n", pseudo_r2_val))
}

if (best_model == "lgbm_twd") {
  # Pseudo-R2 for Tweedie models
  pseudo_r2_val <- compute_pseudo_r2(test_df$strike_count, test_pred, family = "tweedie")
  supplementary_metrics <- bind_rows(supplementary_metrics,
    tibble(.metric = "pseudo_r2_tweedie", .estimator = "standard", .estimate = pseudo_r2_val))
  cat(sprintf(" Pseudo-R² (Tweedie deviance): %.4f\n", pseudo_r2_val))
}

if (best_model %in% c("lm_log", "lgbm_log", "xgb_log")) {
  # For log models, report R2 on log scale
  test_pred_log <- predict(final_fit, new_data = test_df)$.pred
  r2_log <- rsq_vec(log1p(test_df$strike_count), test_pred_log)
  supplementary_metrics <- bind_rows(supplementary_metrics,
    tibble(.metric = "r2_log_scale", .estimator = "standard", .estimate = r2_log))
  cat(sprintf(" R² on log scale: %.4f\n", r2_log))
  cat(" Note: Primary metrics (MAE/RMSE/R²) are on original scale using naive expm1() back-transform\n")
  cat(" To use smearing correction, set use_smearing=TRUE in validation section\n")
}

cat("\n Test metrics (original scale):\n"); print(test_tbl)

# Save outputs for later analysis
arrow::write_parquet(tibble(date = test_df$date, .pred = test_pred), sprintf("../../outputs/models/sst_predictions__%s.parquet", selection_metric))

# Combine standard and supplementary metrics for saving
all_test_metrics <- bind_rows(test_tbl, supplementary_metrics)
readr::write_csv(all_test_metrics, sprintf("../../outputs/models/sst_metrics_%s.csv", selection_metric))

# Save the fitted model for variable importance analysis
saveRDS(final_fit, sprintf("../../outputs/models/sst_best_fitted_model_%s.rds", selection_metric))

# Also save model metadata for reference
model_metadata <- tibble(
  best_model = best_model,
  selection_metric = selection_metric,
  validation_rmse = val_summary$rmse[1],
  validation_mae = val_summary$mae[1],
  validation_rsq = val_summary$rsq[1],
  test_rmse = test_tbl$.estimate[test_tbl$.metric == "rmse"],
  test_mae = test_tbl$.estimate[test_tbl$.metric == "mae"],
  test_rsq = test_tbl$.estimate[test_tbl$.metric == "rsq"]
)

# Add supplementary metric to metadata if available
if (nrow(supplementary_metrics) > 0) {
  for (i in seq_len(nrow(supplementary_metrics))) {
    metric_name <- paste0("test_", supplementary_metrics$.metric[i])
    model_metadata[[metric_name]] <- supplementary_metrics$.estimate[i]
  }
}

readr::write_csv(model_metadata, sprintf("../../outputs/models/sst_model_metadata_%s.csv", selection_metric))
```

## Test Set: Zero vs Non-Zero Metrics

```{r}
#| label: test-separate-metrics

cat("\n[Test] Computing zero/non-zero metrics for best model:", best_model, "\n")

# Use already-computed test predictions from final-test chunk
zero_mask <- test_df$strike_count == 0
nonzero_mask <- test_df$strike_count > 0

test_separate_metrics <- bind_rows(
  tibble(
    model = best_model,
    week_type = "zero",
    .metric = c("mae", "rmse"),
    .estimate = c(
      yardstick::mae_vec(test_df$strike_count[zero_mask], test_pred[zero_mask]),
      yardstick::rmse_vec(test_df$strike_count[zero_mask], test_pred[zero_mask])
    )
  ),
  tibble(
    model = best_model,
    week_type = "non_zero",
    .metric = c("mae", "rmse", "rsq"),
    .estimate = c(
      yardstick::mae_vec(test_df$strike_count[nonzero_mask], test_pred[nonzero_mask]),
      yardstick::rmse_vec(test_df$strike_count[nonzero_mask], test_pred[nonzero_mask]),
      yardstick::rsq_vec(test_df$strike_count[nonzero_mask], test_pred[nonzero_mask])
    )
  )
)

cat("\n[Test] Zero/Non-zero metrics on held-out test set:\n")
print(test_separate_metrics)

readr::write_csv(test_separate_metrics, sprintf("../../outputs/models/sst_test_separate_metrics_%s.csv", selection_metric))
```

## Feature Importance For Best Model (if applicable)

```{r}
#| label: feature-importance

# Extract native importance from tree-based models only
if (best_model %in% c("lgbm_poi", "lgbm_log", "lgbm_twd", "xgb_poi", "xgb_log")) {
  cat(" Extracting native feature importance from", best_model, "\n")
  
  importance_df <- vip::vi_model(final_fit) |>
    arrange(desc(Importance)) |>
    mutate(model = best_model)
  
  # Save importance scores
  arrow::write_parquet(importance_df, sprintf("../../outputs/models/sst_feature_importance_%s.parquet", selection_metric))
  
  # Print top 10
  cat("\n Top 10 most important features:\n")
  print(head(importance_df, 10))
  
} else {
  cat(" Best model is", best_model, "- native importance not available (linear model)\n")
  cat(" Use permutation importance or coefficient analysis for linear models\n")
}
```

## Naive Baseline Comparisons: Validation Set

```{r}
#| label: naive-baselines-validation

cat("\n[Naive Baselines] Computing validation set baseline predictions...\n")

# Helper to compute baseline predictions per fold
compute_baseline_fold <- function(split, baseline_type) {
  analysis_data <- rsample::analysis(split)
  assessment_data <- rsample::assessment(split)
  
  if (baseline_type == "mean") {
    pred <- rep(mean(analysis_data$strike_count, na.rm = TRUE), nrow(assessment_data))
  } else if (baseline_type == "median") {
    pred <- rep(median(analysis_data$strike_count, na.rm = TRUE), nrow(assessment_data))
  } else if (baseline_type == "persistence") {
    pred <- assessment_data$strike_count_lag1
  } else if (baseline_type == "zero") {
    pred <- rep(0, nrow(assessment_data))
  } else if (baseline_type == "rolling_4wk") {
    pred <- assessment_data$rolling_avg_4wk
  } else if (baseline_type == "rolling_8wk") {
    pred <- assessment_data$rolling_avg_8wk
  }
  
  # Clean up predictions (cap at 0, handle NAs)
  pred <- pmax(pred, 0, na.rm = TRUE)
  pred[!is.finite(pred)] <- 0
  
  tibble(
    truth = assessment_data$strike_count,
    .pred = pred,
    baseline_type = baseline_type
  )
}

# Compute all baseline predictions across validation folds
baseline_types <- c("mean", "median", "persistence", "zero", "rolling_4wk", "rolling_8wk")
naive_val_predictions <- list()

for (baseline in baseline_types) {
  cat("[Baseline] Computing", baseline, "predictions across validation folds...\n")
  
  # Get predictions for all folds
  fold_preds <- purrr::map_dfr(val_resamples$splits, ~ compute_baseline_fold(.x, baseline))
  
  # Compute pooled metrics (concatenated across all folds)
  pooled_metrics <- tibble(
    model = paste0("naive_", baseline),
    mae = yardstick::mae_vec(fold_preds$truth, fold_preds$.pred),
    rmse = yardstick::rmse_vec(fold_preds$truth, fold_preds$.pred),
    rsq = yardstick::rsq_vec(fold_preds$truth, fold_preds$.pred)
  )
  
  naive_val_predictions[[baseline]] <- pooled_metrics
}

# Combine all baseline results
naive_val_summary <- bind_rows(naive_val_predictions) |>
  arrange(rmse)  # Rank by RMSE by default

cat("\n[Naive Baselines] Validation set results (ranked by RMSE):\n")
print(naive_val_summary)

# Save validation baseline results (metric-agnostic filename)
readr::write_csv(naive_val_summary, "../../outputs/models/sst_naive_baselines_validation.csv")

# Save validation baseline results only (comparison tables will be created later)
cat("\n[Naive Baselines] Validation baselines saved. Comparison tables will be created during writeup.\n")
```

## Naive Baseline Comparisons: Test Set

```{r}
#| label: naive-baselines-test

cat("\n[Naive Baselines] Computing test set baseline predictions...\n")

# Compute baseline statistics from full pretest data
train_mean <- mean(pretest_df$strike_count, na.rm = TRUE)
train_median <- median(pretest_df$strike_count, na.rm = TRUE)

# Test set baselines
naive_test_results <- list()

# 1. Mean baseline
naive_test_results$mean <- tibble(
  model = "naive_mean",
  mae = yardstick::mae_vec(test_df$strike_count, rep(train_mean, nrow(test_df))),
  rmse = yardstick::rmse_vec(test_df$strike_count, rep(train_mean, nrow(test_df))),
  rsq = yardstick::rsq_vec(test_df$strike_count, rep(train_mean, nrow(test_df)))
)

# 2. Median baseline
naive_test_results$median <- tibble(
  model = "naive_median",
  mae = yardstick::mae_vec(test_df$strike_count, rep(train_median, nrow(test_df))),
  rmse = yardstick::rmse_vec(test_df$strike_count, rep(train_median, nrow(test_df))),
  rsq = yardstick::rsq_vec(test_df$strike_count, rep(train_median, nrow(test_df)))
)

# 3. Persistence (lag1)
naive_test_results$persistence <- tibble(
  model = "naive_persistence",
  mae = yardstick::mae_vec(test_df$strike_count, test_df$strike_count_lag1),
  rmse = yardstick::rmse_vec(test_df$strike_count, test_df$strike_count_lag1),
  rsq = yardstick::rsq_vec(test_df$strike_count, test_df$strike_count_lag1)
)

# 4. Zero baseline
naive_test_results$zero <- tibble(
  model = "naive_zero",
  mae = yardstick::mae_vec(test_df$strike_count, rep(0, nrow(test_df))),
  rmse = yardstick::rmse_vec(test_df$strike_count, rep(0, nrow(test_df))),
  rsq = NA_real_  # undefined for constant prediction
)

# 5. Rolling 4-week average
naive_test_results$rolling_4wk <- tibble(
  model = "naive_rolling_4wk",
  mae = yardstick::mae_vec(test_df$strike_count, test_df$rolling_avg_4wk),
  rmse = yardstick::rmse_vec(test_df$strike_count, test_df$rolling_avg_4wk),
  rsq = yardstick::rsq_vec(test_df$strike_count, test_df$rolling_avg_4wk)
)

# 6. Rolling 8-week average
naive_test_results$rolling_8wk <- tibble(
  model = "naive_rolling_8wk",
  mae = yardstick::mae_vec(test_df$strike_count, test_df$rolling_avg_8wk),
  rmse = yardstick::rmse_vec(test_df$strike_count, test_df$rolling_avg_8wk),
  rsq = yardstick::rsq_vec(test_df$strike_count, test_df$rolling_avg_8wk)
)

# Combine all test baseline results
naive_test_summary <- bind_rows(naive_test_results) |>
  arrange(rmse)  # Rank by RMSE by default

cat("\n[Naive Baselines] Test set results (ranked by RMSE):\n")
print(naive_test_summary)

# Save test baseline results (metric-agnostic filename)
readr::write_csv(naive_test_summary, "../../outputs/models/sst_naive_baselines_test.csv")

# Save test baseline results only (comparison tables will be created later)
cat("\n[Naive Baselines] Test baselines saved. Comparison tables will be created during writeup.\n")

cat("\n[Naive Baselines] All baseline comparisons complete!\n")
```


