---
title: "Country Filter Exploration"
format: html
---

## Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(lubridate)
library(arrow)
library(janitor)
```

## Load data

```{r}
#| label: load-data

adm_full <- read_parquet("../../data/analysis/adm_week_full.parquet") |>
  clean_names() |>
  mutate(date = as.Date(date)) |>
  select(date, strike_count, action_geo_country_code, action_geo_adm1_code) |>
  arrange(date)

glimpse(adm_full)
```

## Pretest/Test split (match modeling scripts)

```{r}
#| label: splits

all_weeks <- sort(unique(adm_full$date))
test_weeks <- tail(all_weeks, 52)

pretest_df <- adm_full |>
  filter(date < min(test_weeks))
test_df <- adm_full |>
  filter(date %in% test_weeks)

cat("Pretest rows:", nrow(pretest_df), " Test rows:", nrow(test_df), "\n")
```

## Drop impact by leak-safe country eligibility rules

```{r}
#| label: country-filter-impact

# Compute summary for a given threshold rule
summarize_drop_impact <- function(thr_label, thr_value) {
  thr_value_num <- suppressWarnings(as.numeric(thr_value))
  if (is.na(thr_value_num)) {
    # ever-positive rule (binary)
    eligible <- pretest_df |>
      group_by(action_geo_country_code) |>
      summarise(any_pos = any(strike_count > 0, na.rm = TRUE), .groups = "drop") |>
      filter(any_pos) |>
      pull(action_geo_country_code)
  } else {
    # rate threshold (share of positive weeks in pretest)
    eligible <- pretest_df |>
      group_by(action_geo_country_code) |>
      summarise(pos_rate = mean(strike_count > 0, na.rm = TRUE), .groups = "drop") |>
      filter(pos_rate >= thr_value_num) |>
      pull(action_geo_country_code)
  }

  pre_f <- pretest_df |> filter(action_geo_country_code %in% eligible)
  test_f <- test_df    |> filter(action_geo_country_code %in% eligible)

  tibble(
    rule = thr_label,
    countries_pretest_total = n_distinct(pretest_df$action_geo_country_code),
    countries_pretest_kept  = n_distinct(pre_f$action_geo_country_code),
    countries_pretest_drop  = countries_pretest_total - countries_pretest_kept,
    countries_test_total    = n_distinct(test_df$action_geo_country_code),
    countries_test_kept     = n_distinct(test_f$action_geo_country_code),
    countries_test_drop     = countries_test_total - countries_test_kept,
    rows_pretest_before     = nrow(pretest_df),
    rows_pretest_after      = nrow(pre_f),
    pretest_drop_pct        = 1 - rows_pretest_after / pmax(rows_pretest_before, 1),
    rows_test_before        = nrow(test_df),
    rows_test_after         = nrow(test_f),
    test_drop_pct           = 1 - rows_test_after / pmax(rows_test_before, 1),
    adm1_pretest_total      = n_distinct(pretest_df$action_geo_adm1_code),
    adm1_pretest_kept       = n_distinct(pre_f$action_geo_adm1_code),
    adm1_pretest_drop       = adm1_pretest_total - adm1_pretest_kept,
    adm1_test_total         = n_distinct(test_df$action_geo_adm1_code),
    adm1_test_kept          = n_distinct(test_f$action_geo_adm1_code),
    adm1_test_drop          = adm1_test_total - adm1_test_kept
  )
}

thr_list <- list(
  ever = NA_real_,
  rate_0_001 = 0.001,
  rate_0_002 = 0.002,
  rate_0_005 = 0.005
)

drop_summary <- purrr::imap_dfr(thr_list, ~ summarize_drop_impact(.y, .x)) |>
  mutate(
    pretest_drop_pct = scales::percent(pretest_drop_pct, accuracy = 0.1),
    test_drop_pct    = scales::percent(test_drop_pct, accuracy = 0.1)
  )

drop_summary |>
  arrange(desc(rows_test_before - rows_test_after)) |>
  print(n = Inf)
```

## Row drops (explicit counts)

```{r}
#| label: row-drops-counts

drop_summary |>
  transmute(
    rule,
    rows_pretest_drop = rows_pretest_before - rows_pretest_after,
    pretest_drop_pct,
    rows_test_drop = rows_test_before - rows_test_after,
    test_drop_pct
  ) |>
  arrange(desc(rows_test_drop)) |>
  print(n = Inf)
```

## Country contributions to row drops (accounts for different ADM1 counts)

```{r}
#| label: country-drop-breakdown

country_rows_pre <- pretest_df |>
  count(action_geo_country_code, name = "rows_pretest")
country_rows_test <- test_df |>
  count(action_geo_country_code, name = "rows_test")
adm1_per_country <- pretest_df |>
  distinct(action_geo_country_code, action_geo_adm1_code) |>
  count(action_geo_country_code, name = "adm1_count_pretest")

country_drop_table <- function(thr_label, thr_value) {
  thr_value_num <- suppressWarnings(as.numeric(thr_value))
  if (is.na(thr_value_num)) {
    elig <- pretest_df |>
      group_by(action_geo_country_code) |>
      summarise(any_pos = any(strike_count > 0, na.rm = TRUE), .groups = "drop") |>
      filter(any_pos) |>
      pull(action_geo_country_code)
  } else {
    elig <- pretest_df |>
      group_by(action_geo_country_code) |>
      summarise(pos_rate = mean(strike_count > 0, na.rm = TRUE), .groups = "drop") |>
      filter(pos_rate >= thr_value_num) |>
      pull(action_geo_country_code)
  }

  country_rows_pre |>
    full_join(country_rows_test, by = "action_geo_country_code") |>
    left_join(adm1_per_country, by = "action_geo_country_code") |>
    mutate(
      rows_pretest = coalesce(rows_pretest, 0L),
      rows_test = coalesce(rows_test, 0L),
      adm1_count_pretest = coalesce(adm1_count_pretest, 0L),
      kept = action_geo_country_code %in% elig,
      rows_pretest_dropped = if_else(kept, 0L, rows_pretest),
      rows_test_dropped = if_else(kept, 0L, rows_test),
      rule = thr_label
    ) |>
    arrange(desc(rows_test_dropped))
}

country_drops <- purrr::imap_dfr(thr_list, ~ country_drop_table(.y, .x))

# Per-rule totals (kept/dropped counts and rows)
rule_totals <- country_drops |>
  group_by(rule) |>
  summarise(
    countries_kept = sum(kept),
    countries_dropped = sum(!kept),
    rows_pretest_dropped = sum(rows_pretest_dropped),
    rows_test_dropped = sum(rows_test_dropped),
    .groups = "drop"
  )
print(rule_totals)

# Top dropped countries per rule (only those actually dropped)
top_dropped <- country_drops |>
  filter(rows_test_dropped > 0) |>
  group_by(rule) |>
  slice_max(order_by = rows_test_dropped, n = 15, with_ties = FALSE) |>
  ungroup() |>
  arrange(rule, desc(rows_test_dropped))

if (nrow(top_dropped) == 0) {
  cat("No countries dropped under any rule.\n")
} else {
  print(top_dropped, n = Inf)
}

```

## ADM1-level filter impact (ever-positive / min positives / min rate)

```{r}
#| label: adm1-filter-impact

# Precompute ADM1-level stats on pretest
adm1_stats_pre <- pretest_df |>
  group_by(action_geo_adm1_code, action_geo_country_code) |>
  summarise(
    weeks = n(),
    pos_weeks = sum(strike_count > 0, na.rm = TRUE),
    pos_rate = pos_weeks / weeks,
    .groups = "drop"
  )

adm1_rows_pre <- pretest_df |>
  count(action_geo_adm1_code, action_geo_country_code, name = "rows_pretest")
adm1_rows_test <- test_df |>
  count(action_geo_adm1_code, action_geo_country_code, name = "rows_test")

# Define ADM1 rules
adm1_rules <- list(
  ever = list(min_pos = 1L, min_rate = NA_real_),
  min_pos_2 = list(min_pos = 2L, min_rate = NA_real_),
  min_pos_4 = list(min_pos = 4L, min_rate = NA_real_),
  rate_0_001 = list(min_pos = NA_integer_, min_rate = 0.001),
  rate_0_002 = list(min_pos = NA_integer_, min_rate = 0.002),
  rate_0_005 = list(min_pos = NA_integer_, min_rate = 0.005)
)

eligible_adm1 <- function(rule) {
  if (!is.na(rule$min_pos)) {
    adm1_stats_pre |>
      filter(pos_weeks >= rule$min_pos) |>
      pull(action_geo_adm1_code)
  } else if (!is.na(rule$min_rate)) {
    adm1_stats_pre |>
      filter(pos_rate >= rule$min_rate) |>
      pull(action_geo_adm1_code)
  } else {
    adm1_stats_pre |>
      filter(pos_weeks > 0) |>
      pull(action_geo_adm1_code)
  }
}

summarize_adm1_drop_impact <- function(thr_label, rule) {
  elig <- eligible_adm1(rule)

  pre_f <- pretest_df |> filter(action_geo_adm1_code %in% elig)
  test_f <- test_df    |> filter(action_geo_adm1_code %in% elig)

  tibble(
    rule = thr_label,
    adm1_pretest_total = n_distinct(pretest_df$action_geo_adm1_code),
    adm1_pretest_kept  = n_distinct(pre_f$action_geo_adm1_code),
    adm1_pretest_drop  = adm1_pretest_total - adm1_pretest_kept,
    rows_pretest_before = nrow(pretest_df),
    rows_pretest_after  = nrow(pre_f),
    pretest_drop_pct     = 1 - rows_pretest_after / pmax(rows_pretest_before, 1),
    rows_test_before     = nrow(test_df),
    rows_test_after      = nrow(test_f),
    test_drop_pct        = 1 - rows_test_after / pmax(rows_test_before, 1)
  )
}

drop_summary_adm1 <- purrr::imap_dfr(adm1_rules, ~ summarize_adm1_drop_impact(.y, .x)) |>
  mutate(
    pretest_drop_pct = scales::percent(pretest_drop_pct, accuracy = 0.1),
    test_drop_pct    = scales::percent(test_drop_pct, accuracy = 0.1)
  )

drop_summary_adm1 |>
  arrange(desc(rows_test_before - rows_test_after)) |>
  print(n = Inf)
```

## ADM1s dropped per rule and country totals from ADM1 filtering

```{r}
#| label: adm1-dropped-breakdown

adm1_drop_table <- function(thr_label, rule) {
  elig <- eligible_adm1(rule)

  adm1_rows_pre |>
    full_join(adm1_rows_test, by = c("action_geo_adm1_code","action_geo_country_code")) |>
    mutate(
      rows_pretest = coalesce(rows_pretest, 0L),
      rows_test = coalesce(rows_test, 0L),
      kept = action_geo_adm1_code %in% elig,
      rows_pretest_dropped = if_else(kept, 0L, rows_pretest),
      rows_test_dropped = if_else(kept, 0L, rows_test),
      rule = thr_label
    ) |>
    arrange(desc(rows_test_dropped))
}

adm1_drops <- purrr::imap_dfr(adm1_rules, ~ adm1_drop_table(.y, .x))

# Top 20 ADM1 by test rows dropped per rule
top_adm1_dropped <- adm1_drops |>
  filter(rows_test_dropped > 0) |>
  group_by(rule) |>
  slice_max(order_by = rows_test_dropped, n = 20, with_ties = FALSE) |>
  ungroup() |>
  arrange(rule, desc(rows_test_dropped))

if (nrow(top_adm1_dropped) == 0) {
  cat("No ADM1 regions dropped under any rule.\n")
} else {
  print(top_adm1_dropped, n = Inf)
}

# Country totals resulting from ADM1 filtering
adm1_country_totals <- adm1_drops |>
  group_by(rule, action_geo_country_code) |>
  summarise(
    rows_pretest_dropped = sum(rows_pretest_dropped),
    rows_test_dropped = sum(rows_test_dropped),
    adm1_dropped = sum(rows_test_dropped > 0),
    .groups = "drop"
  ) |>
  arrange(rule, desc(rows_test_dropped))

print(adm1_country_totals, n = 50)
```

## Optional: inspect countries dropped under each rule

```{r}
#| label: inspect-dropped
#| eval: false

inspect_dropped <- function(thr_value) {
  if (is.na(thr_value)) {
    elig <- pretest_df |>
      group_by(action_geo_country_code) |>
      summarise(any_pos = any(strike_count > 0), .groups = "drop") |>
      filter(any_pos) |>
      pull(action_geo_country_code)
  } else {
    elig <- pretest_df |>
      group_by(action_geo_country_code) |>
      summarise(pos_rate = mean(strike_count > 0), .groups = "drop") |>
      filter(pos_rate >= thr_value) |>
      pull(action_geo_country_code)
  }
  setdiff(unique(adm_full$action_geo_country_code), elig)
}

inspect_dropped(NA)      # ever-positive rule
# inspect_dropped(0.002) # rate >= 0.2%
```


