---
title: "The Spatial Diffusion of Labor Strikes"
subtitle: "Evidence from Global Subnational Event Data"
author: "Emmanuel Teitelbaum"
institute: "George Washington University"
date: today
date-format: long
footer: "APSA Labor Group CPLW · March 2026"
logo: images/esia-logo.png
format:
  revealjs:
    theme: [simple, light-gw.scss]
    transition: fade
    slide-number: true
    chalkboard: true
    width: 1280
    height: 720
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(here)
library(tidyverse)
library(fixest)
library(modelsummary)
library(knitr)
library(kableExtra)
library(scales)

source(here("papers/contagion/helpers/figures.R"))
source(here("papers/contagion/helpers/tables.R"))

# Shared modelsummary options
ms_stars  <- c("*" = 0.1, "**" = 0.05, "***" = 0.01)
ms_gof    <- tribble(
  ~raw,          ~clean,      ~fmt,
  "nobs",        "N",          0,
  "r.squared",   "R²",         3,
  "r2.within",   "Within R²",  3
)
```

# The Puzzle

## Strike waves are real...

```{r}
#| label: fig-puzzle
#| fig-width: 10
#| fig-height: 4.5

fig_1
```

::: {.fragment}
Regions with a striking neighbor are **3.5× more likely** to strike. Is this geographic contagion — or just national common shocks?
:::

# Theory

## Two mechanisms, one pattern {.smaller}

:::: {.columns}

::: {.column width="48%"}
**Geographic contagion**

 Workers observe a neighboring region striking and update beliefs about the feasibility and costs of collective action, lowering the threshold for action nearby. 

<br>

*Relational mechanism: transmission through direct ties.*
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
**National contagion**

Union federations, national policy, and media simultaneously shift strike propensity across all regions — producing clustering without any direct transmission.

<br>

*Non-relational mechanism: independent response to common stimuli.*
:::

::::

<br>

::: {.fragment}
**Both produce identical observable patterns.** Historical (e.g. Shorter & Tilly 1974; Franzosi 1995) and economic (e.g. Ashenfelter & Johnson 1969; Hibbs 1976) accounts both lean national, but no prior study has identified the two mechanisms causally at subnational scale.        
:::

# Data

## Filtering raw records to validated strikes

```{r}
#| label: tbl-pipeline

pipeline <- tibble(
  Step = c(
    "BigQuery download",
    "Geographic filter",
    "High-confidence filter",
    "Keyword theme filter",
    "LLM content classifier"
  ),
  Description = c(
    "CAMEO code 143, valid coordinates",
    "Valid subnational region code assigned",
    "At least one high-confidence mention",
    "Article tagged \"STRIKE\" in topic metadata",
    "Article describes a labor strike (GPT-4o mini)"
  ),
  Records = c("492,500", "328,982", "134,141", "30,899", "11,503")
)

kable(pipeline, align = c("l", "l", "r")) |>
  kable_styling(font_size = 22, full_width = TRUE) |>
  row_spec(5, bold = TRUE)
```

## Global coverage

```{r}
#| label: fig-map
#| fig-width: 10
#| fig-height: 4.5

fig_2
```

::: {style="font-size: 0.75em; color: grey;"}
568 ADM1 regions, 74 countries, 2015–2025. Clustering reflects genuine activity and GDELT's stronger coverage of English-language and romanized-script sources.
:::

# Identification

## Baseline

**Test for geographic contagion**

$$Y_{it} = \beta \cdot \text{NeighborStrike}_{i,t-1:t-2} + \alpha_i + \gamma_{c \times t} + \varepsilon_{it}$$

Country-by-week FEs $(\gamma_{c \times t})$ absorb *every* national-level shock. (No need to specify which factors matter.)

<br>

::: {style="font-size: 0.8em; color: grey; margin-top: 0.5em;"}
Estimated with LPM via `feols`; standard errors clustered at ADM1 level.
:::

## Decomposition 

**Test for separate national and geographic channels**

$$\begin{aligned}
Y_{it} =\, &\beta_1 \cdot \text{NeighborStrike}_{i,t-1:t-2} + \beta_2 \cdot \text{CountryShare}_{i,t-1:t-2} \\
            &+ \alpha_i + \gamma_t + \varepsilon_{it}
\end{aligned}$$ 

Replaces country×week FE with an explicit measure of national strike activity. Estimates both channels simultaneously.

<br>

::: {style="font-size: 0.8em; color: grey; margin-top: 0.5em;"}
Estimated with LPM via `feols`; standard errors clustered at ADM1 level.
:::

# Results

## No evidence of geographic contagion

```{r}
#| label: tbl-baseline

modelsummary(
  tbl_baseline,
  coef_map = c(
    "neighbor_strike_within_t1_t2" = "Neighbor strike (t−1, t−2)",
    "neighbor_strike_within_t1_t1" = "Neighbor strike (t−1 only)",
    "neighbor_strike_within_t2_t2" = "Neighbor strike (t−2 only)"
  ),
  gof_map  = ms_gof,
  stars    = ms_stars,
  col.names = c(" ", "(1) Pooled", "(2) t−1", "(3) t−2"),
  output   = "kableExtra"
) |>
  kable_styling(font_size = 20, full_width = FALSE, position = "center") |>
  add_header_above(c(" " = 1, "Country × Week FEs" = 3)) |>
  footnote(general = "ADM1 fixed effects included in all models. SEs clustered at ADM1 level.",
           footnote_as_chunk = TRUE, general_title = "")
```

::: {.fragment}
Raw 3.5× neighbor association **fully disappears** once country-by-week FEs are added.
:::

## National contagion dominates geographic effects

```{r}
#| label: tbl-decomp

modelsummary(
  tbl_decomposition,
  coef_map = c(
    "country_share_t1_t2"          = "Country share (t−1, t−2)",
    "neighbor_strike_within_t1_t2" = "Neighbor strike (t−1, t−2)"
  ),
  gof_map  = ms_gof,
  stars    = ms_stars,
  col.names = c(" ", "(1) Country", "(2) Neighbor", "(3) Both"),
  output   = "kableExtra"
) |>
  kable_styling(font_size = 20, full_width = FALSE, position = "center") |>
  add_header_above(c(" " = 1, "Country + Week FEs" = 3)) |>
  footnote(general = "ADM1 and week fixed effects in all models. SEs clustered at ADM1 level.",
           footnote_as_chunk = TRUE, general_title = "")
```

::: {.fragment}
1 SD increase in country share **more than doubles** strike probability. National contagion explains **63× more** within-unit variance than geographic exposure.
:::

# Conclusion

## Strike contagion is national, not geographic {.smaller}

**Three robust findings:**

1. The raw geographic clustering is real but does not survive a design that absorbs national common shocks

2. National contagion is large: when strikes rise elsewhere in the country, local strike probability rises substantially

3. Institutions shape the *spatial form* of national waves (freedom of expression flattens geographic structure; freedom of association reinforces it)

<br>

**Broader implication:** Geographic proximity is a marker of shared national exposure rather than a conduit of transmission. Future research should focus on what triggers national synchronization, e.g. economic, political, and institutional factors that generate country-wide waves.

# Appendix {visibility="uncounted"}

## Institutions shape the geography of national waves {visibility="uncounted"}

```{r}
#| label: tbl-moderation

modelsummary(
  tbl_moderation_ext,
  coef_map = c(
    "country_share_t1_t2"                     = "Country share",
    "neighbor_strike_within_t1_t2"            = "Neighbor strike",
    "neighbor_strike_within_t1_t2:freexp_z"   = "Neighbor × Expression",
    "neighbor_strike_within_t1_t2:frassoc_z"  = "Neighbor × Association",
    "neighbor_strike_within_t1_t2:repression_z" = "Neighbor × Repression",
    "country_share_t1_t2:freexp_z"            = "Country share × Expression",
    "country_share_t1_t2:frassoc_z"           = "Country share × Association",
    "country_share_t1_t2:repression_z"        = "Country share × Repression"
  ),
  gof_map  = ms_gof,
  stars    = ms_stars,
  col.names = c(" ", "(1) Expression", "(2) Association", "(3) Repression", "(4) Joint"),
  output   = "kableExtra"
) |>
  kable_styling(font_size = 16, full_width = FALSE, position = "center") |>
  footnote(general = "ADM1 and week FEs. All V-Dem indicators standardized. SEs clustered at ADM1 level.",
           footnote_as_chunk = TRUE, general_title = "")
```

## Robustness {visibility="uncounted"}

- **Bias-corrected logit** (`alpaca`): substantive conclusions unchanged
- **Two-way clustering** (ADM1 + country-week): null geographic result holds
- **Validated-location subsample** (confirmed ADM1 geocodes only): null geographic \& large national both replicate
- **Cross-border neighbors**: null across all specifications; strike contagion bounded by national frameworks
