---
title: "Contagion Panel: Diagnostics"
format:
  html:
    code-fold: true
    toc: true
    fig-width: 10
    fig-height: 6
---

Exploratory diagnostics for `data/analysis/contagion_panel.parquet`. This
notebook loads the final analysis panel and visualizes its key properties —
coverage, strike distributions, neighbor exposure, and institutional moderators.

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(lubridate)
library(arrow)
library(sf)
library(scales)
library(here)

conflicted::conflicts_prefer(dplyr::filter)
conflicted::conflicts_prefer(dplyr::lag)
conflicted::conflicts_prefer(lubridate::date)

theme_set(theme_minimal(base_size = 12))

panel <- read_parquet(here("data/analysis/contagion_panel.parquet"))
adm1_sf <- st_read(
  here("data/spatial/gadm_adm1_simplified.gpkg"),
  quiet = TRUE
)

cat("Panel rows:    ", nrow(panel), "\n")
cat("ADM1 units:    ", n_distinct(panel$gid_1), "\n")
cat("Countries:     ", n_distinct(panel$gid_0), "\n")
cat("Weeks:         ", n_distinct(panel$week_start), "\n")
cat("Date range:    ", format(min(panel$week_start)), "–",
                       format(max(panel$week_start)), "\n")
cat("Strike onset %:", round(mean(panel$strike_onset) * 100, 2), "\n")
```

---

## 1. Strike events over time

Global weekly strike onset count across all ADM1 units.

```{r}
#| label: strikes-over-time

panel |>
  group_by(week_start) |>
  summarize(n_striking = sum(strike_onset), .groups = "drop") |>
  ggplot(aes(week_start, n_striking)) +
  geom_line(linewidth = 0.4, color = "steelblue") +
  geom_smooth(method = "loess", span = 0.15, se = FALSE,
              color = "firebrick", linewidth = 0.8) +
  labs(
    title = "Weekly ADM1 strike onsets (global)",
    x = NULL, y = "ADM1 units with ≥1 strike"
  )
```

---

## 2. Strike frequency by country

Top 30 countries by total ADM1-weeks with strike onset.

```{r}
#| label: strikes-by-country

panel |>
  group_by(gid_0) |>
  summarize(total_onset_weeks = sum(strike_onset), .groups = "drop") |>
  slice_max(total_onset_weeks, n = 30) |>
  mutate(gid_0 = fct_reorder(gid_0, total_onset_weeks)) |>
  ggplot(aes(total_onset_weeks, gid_0)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Top 30 countries by ADM1-weeks with strike onset",
    x = "ADM1-weeks with strike", y = NULL
  )
```

---

## 3. Strike frequency by ADM1 unit

Distribution of strike onset rates across ADM1 units (excluding zero-strike
units).

```{r}
#| label: adm1-strike-rate

adm1_rates <- panel |>
  group_by(gid_1, gid_0) |>
  summarize(
    n_weeks       = n(),
    n_onset       = sum(strike_onset),
    onset_rate    = mean(strike_onset),
    .groups = "drop"
  ) |>
  filter(n_onset > 0)

cat("ADM1 units with ≥1 strike onset:", nrow(adm1_rates), "/",
    n_distinct(panel$gid_1), "\n")

ggplot(adm1_rates, aes(onset_rate)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  scale_x_continuous(labels = percent) +
  labs(
    title = "Distribution of strike onset rates across ADM1 units",
    subtitle = "Units with ≥1 strike onset only",
    x = "Strike onset rate (fraction of weeks)", y = "ADM1 units"
  )
```

---

## 4. Neighbor strike exposure

Distribution of the preregistered treatment variable
(`neighbor_strike_within_t1_t2`).

```{r}
#| label: exposure-dist

panel |>
  filter(!is.na(neighbor_strike_within_t1_t2)) |>
  count(neighbor_strike_within_t1_t2) |>
  mutate(
    pct   = n / sum(n),
    label = c("No neighbor strike", "Neighbor strike")
  ) |>
  ggplot(aes(label, pct)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Neighbor strike exposure (within-country, t-1 to t-2)",
    x = NULL, y = "Share of ADM1-weeks"
  )
```

```{r}
#| label: exposure-by-outcome

# Key descriptive: strike rate conditional on neighbor exposure
panel |>
  filter(!is.na(neighbor_strike_within_t1_t2)) |>
  group_by(neighbor_strike_within_t1_t2) |>
  summarize(
    n            = n(),
    strike_rate  = mean(strike_onset),
    .groups      = "drop"
  ) |>
  mutate(
    label       = c("No neighbor strike", "Neighbor strike"),
    pct         = percent(strike_rate, accuracy = 0.1)
  ) |>
  ggplot(aes(label, strike_rate)) +
  geom_col(fill = c("steelblue", "firebrick")) +
  geom_text(aes(label = pct), vjust = -0.4, size = 4) +
  scale_y_continuous(labels = percent, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Strike onset rate by neighbor exposure",
    subtitle = "Raw (unconditional) association",
    x = NULL, y = "Strike onset rate"
  )
```

---

## 5. Geo quality flag

Breakdown of validated vs. unvalidated events contributing to strike onsets.

```{r}
#| label: geo-quality

panel |>
  filter(strike_onset == 1) |>
  count(geo_quality) |>
  mutate(pct = percent(n / sum(n), accuracy = 0.1)) |>
  knitr::kable(caption = "Strike onset weeks by geo_quality")
```

---

## 6. V-Dem institutional moderators

```{r}
#| label: vdem-dist

panel |>
  distinct(gid_0, week_start, freexp, frassoc, repression) |>
  select(freexp, frassoc, repression) |>
  pivot_longer(everything(), names_to = "variable",
               names_ptypes = list(variable = factor())) |>
  mutate(variable = recode(variable,
    freexp     = "Freedom of expression",
    frassoc    = "Freedom of association",
    repression = "State repression"
  )) |>
  ggplot(aes(value)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  facet_wrap(~variable, scales = "free_y") +
  labs(
    title = "V-Dem institutional moderator distributions",
    subtitle = "Country-year level (repeated across weeks within year)",
    x = "V-Dem index value", y = "Country-years"
  )
```

```{r}
#| label: vdem-summary

panel |>
  distinct(gid_0, week_start, freexp, frassoc, repression) |>
  select(freexp, frassoc, repression) |>
  pivot_longer(everything(), names_to = "variable") |>
  group_by(variable) |>
  summarize(
    n       = sum(!is.na(value)),
    mean    = round(mean(value, na.rm = TRUE), 3),
    sd      = round(sd(value, na.rm = TRUE), 3),
    p10     = round(quantile(value, .10, na.rm = TRUE), 3),
    p50     = round(quantile(value, .50, na.rm = TRUE), 3),
    p90     = round(quantile(value, .90, na.rm = TRUE), 3),
    missing = sum(is.na(value)),
    .groups = "drop"
  ) |>
  knitr::kable(caption = "V-Dem moderator summary statistics")
```

---

## 7. Coverage map

Which ADM1 units are in the panel, and how many strike onsets did each have?

```{r}
#| label: coverage-map
#| fig-height: 7

strike_counts <- panel |>
  group_by(gid_1) |>
  summarize(n_onset = sum(strike_onset), .groups = "drop")

map_data <- adm1_sf |>
  left_join(strike_counts, by = c("GID_1" = "gid_1")) |>
  mutate(
    status = case_when(
      is.na(n_onset) ~ "Not in panel",
      n_onset == 0   ~ "In panel, no strikes",
      TRUE           ~ "In panel, ≥1 strike"
    ),
    n_onset_log = log1p(n_onset)
  )

ggplot(map_data) +
  geom_sf(aes(fill = n_onset_log), linewidth = 0.05, color = "white") +
  geom_sf(
    data = map_data |> filter(is.na(n_onset)),
    fill = "#d9d9d9", linewidth = 0.05, color = "white"
  ) +
  scale_fill_gradient(
    low = "#c6dbef", high = "#08306b",
    na.value = "#d9d9d9",
    name = "Weeks",
    labels = \(x) round(expm1(x)),
    breaks = log1p(c(2, 6, 19, 54, 147))
  ) +
  labs(
    title    = "Strike Prevalence Across Subnational Regions, 2015–2025",
    subtitle = "Number of weeks with recorded strike activity.",
    caption  = "Strike data derived from GDELT; boundaries from GADM."
  ) +
  theme_minimal() +
  theme(
    legend.position  = "right",
    axis.text        = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.2)
  )
```

```{r}
#| label: isolates-map
#| fig-height: 7

# Highlight isolates (no within-country neighbors)
adj <- readRDS(here("data/spatial/gadm_adjacency.rds"))
isolates <- adj$summary |> filter(n_neighbors_within == 0)

map_isolates <- adm1_sf |>
  filter(GID_1 %in% unique(panel$gid_1)) |>
  mutate(is_isolate = GID_1 %in% isolates$gid_1)

ggplot(map_isolates) +
  geom_sf(aes(fill = is_isolate), linewidth = 0.05, color = "white") +
  scale_fill_manual(
    values = c("FALSE" = "#c6dbef", "TRUE" = "firebrick"),
    labels = c("FALSE" = "Has neighbors", "TRUE" = "Isolate (no within-country neighbor)"),
    name = NULL
  ) +
  labs(
    title = "ADM1 isolates: units with no within-country neighbors",
    subtitle = "Isolates receive NA exposure and are excluded from main specification"
  ) +
  theme_minimal() +
  theme(
    legend.position  = "bottom",
    axis.text        = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.2)
  )
```
